### 2011-12-22 paper/paper.tex
Begin working on the paper

\documentclass[11pt,a4paper,english,twocolumn]{article}
%\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage{blindtext}
\usepackage{hyperref}

\title{\textbf{Slovene NLTK Tagger}}
\author{
Niko Colnerič \\
\footnotesize Faculty of Computer and \\
\footnotesize Information Science \\
\footnotesize University of Ljubljana \\
\and
Nejc Banič \\
\footnotesize Faculty of Computer and \\
\footnotesize Information Science \\
\footnotesize University of Ljubljana \\
}

\begin{document}
\maketitle
\thispagestyle{empty}
%\pagebreak

\section*{Abstract}
\textbf{TODO - NIKICC}

\section{Introduction}
\subsection{Trigram tagger}
\subsection{Brill tagger}
\subsection{Naive Bayes tagger}
\section{Corpus}
\section{NLTK Trainer}
\section{Implementation}
\section{Conclusion}

%\tableofcontents\newpage
\blinddocument

%\thispagestyle{empty}
\begin{thebibliography}{99}
\bibitem{JOS} \url{http://nl.ijs.si/jos/}

\end{thebibliography}
\end{document}

### 2011-12-22 example.py
Modified example.py

@@ -71,8 +71,9 @@ _t1 = time.time()
 
 
 #uText is input text
-uText = u'Lep je dan, vse diši že po pomladi!'
-uText = u'Živé naj vsi naródi, ki hrepené dočakat dan, da, koder sonce hodi, prepir iz svéta bo pregnan, ko rojak prost bo vsak, ne vrag, le sosed bo mejak.'
+#uText = u'Lep je dan, vse diši že po pomladi!'
+#uText = u'Živé naj vsi naródi, ki hrepené dočakat dan, da, koder sonce hodi, prepir iz svéta bo pregnan, ko rojak prost bo vsak, ne vrag, le sosed bo mejak.'
+uText = u'Potem se je obrnil proti nam in skozi solze izoblikoval najlepši nasmeh, kar sem jih videl v življenju.'
 
 
 _t2 = time.time()

### 2011-12-29 paper/paper.tex
Some work on the paper, modified generation script. The fraction should be almost 1 not almost 0 and the evaluation woks just fine

@@
 
-\section*{Abstract}
-\textbf{TODO - NIKICC}
-
-\section{Introduction}
+\section*{Abstract - NIKICC}
+This paper describes the process of building Slovene tagger for use in NLTK.  
+\section{Introduction - NIKICC}
+Tagger processes a sequence of words and attaches a part of speech tag to each word.
+We implemented tagger for Slovene language in NLTK.
+NLTK stands for Natural Language Toolkit and is a library for use in Python (currently for version below 3, we used 2.7).
+NLTK already has some taggers for English and some other languages.
+Our goal was to implement NLTK tagger for Slovene language.
+The resault of this project depends on two other larger projects.
+First is \textit{JOS}\cite{JOS}(in Slovene it stands for: "Jezikoslovno Označevanje Slovenskega jezika"), from which we used the corpuses (large and structured set of texts).
+Second, \textit{nltk-trainer}\cite{nltk-trainer}, is an open-source project hosting on Github, which was used for actuall training of the tagger on Slovene sentences.
+Our work based mainly on putting all this pieces together into a working Slovene NLTK tagger.
+\section{JOS Corpus}
+\textit{TODO - BANIČ}
+\section{Nltk-trainer - NIKICC}
+Nltk-trainer\cite{nltk-trainer} is project, whose author is Jacob Perkins, from which we took the code for building the tagger.
+We almost exclusively used the script \textit{train\_tagger.py}, which has the ability to generate NLTK taggers.
+Its mandatory argument is corpus in .pos format.
+Other optional arguments can choose wheather to generate a Sequential Tagger, a Brill Tagger or a Classifier Based Tagger.
+It also has the possibility to set default tag (the tag for words, for which the tagger doesn't know how to tag) and wheather to evaluate the tagger or not.
+For evaluation of the tagger (section \ref{Evaluation}) this argument was used.
+\section{Taggers description}
 \subsection{Trigram tagger}
+\textit{TODO - NIKICC}
 \subsection{Brill tagger}
+\textit{TODO - NIKICC}
 \subsection{Naive Bayes tagger}
-\section{Corpus}
-\section{NLTK Trainer}
+\textit{TODO - BANIČ}
 \section{Implementation}
+\subsection{Corpus transformation}
+\textit{TODO - BANIČ}
+\subsection{Training the tagger}
+\textit{TODO - NIKICC}
+\subsection{Usage with NLTK}
+\textit{TODO - BANIČ}
+\subsection{Encoding problems}
+\textit{TODO - BANIČ}
+\section{Evaluation}\label{Evaluation}
+\textit{TODO - NIKICC, BANIČ}
 \section{Conclusion}
+\textit{TODO - BANIČ}
 
 %\tableofcontents\newpage
-\blinddocument
+%\blinddocument
 
@@
 \bibitem{JOS} \url{http://nl.ijs.si/jos/}
+\bibitem{nltk-trainer} \url{https://github.com/japerk/nltk-trainer}
 

### 2011-12-29 generateTaggers.sh

@@ -1,44 +1,36 @@
 # For evaluating the taggers after construction remove "--no-eval" argument.
-# By defult this evaluate on the whole corpus, which might take a lot of time.
-# To evaluate tagger on only a fraction of a corpus use argument "--fraction FRACTION"
+# By defult this evaluate on the whole corpus.
+# To evaluate tagger on only a fraction of a corpus (which is NOT used for training)
+# use argument "--fraction FRACTION"
 
+#
+# GENERATION
+#
+
+date
 # TrigramTagger without evaluation
 python trainer/train_tagger.py --filename slovene_taggers/TrigramTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --default="-Neznan-" --no-eval ./pos
-
+date
 # BrillTagger without evaluation
 python trainer/train_tagger.py --filename slovene_taggers/BrillTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --default="-Neznan-" --brill --no-eval ./pos
-
+date
 # ClassifierBasedPOSTagger - NaiveBayes without evaluation
 python trainer/train_tagger.py --filename slovene_taggers/NaiveBayes.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --default="-Neznan-" --sequential='' --classifier="NaiveBayes" --no-eval ./pos
-
-# ClassifierBasedPOSTagger - Maxent without evaluation
-python trainer/train_tagger.py --filename slovene_taggers/Maxent.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --default="-Neznan-" --sequential='' --classifier="Maxent" --no-eval ./pos
-
-# ClassifierBasedPOSTagger - NaiveBayes, Maxent without evaluation
-python trainer/train_tagger.py --filename slovene_taggers/NaiveBayesMaxent.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --default="-Neznan-" --sequential='' --classifier=["NaiveBayes","Maxent"] --no-eval ./pos
-
-# ClassifierBasedPOSTagger - DecisionTree without evaluation
[…]

### 2011-12-29 evaluateTaggers.sh
Separated evalation and generation script

@@ -1,0 +1,23 @@
+# For evaluating the taggers after construction remove "--no-eval" argument.
+# By defult this evaluate on the whole corpus.
+# To evaluate tagger on only a fraction of a corpus (which is NOT used for training)
+# use argument "--fraction FRACTION"
+
+#
+# EVALUATION (--fraction 0.99)
+# use 99% for generating the tagger and 1% for evaluation 
+# this is only used for evaluation, for real generation look above
+#
+
+mkdir tmp
+
+date
+# TrigramTagger without evaluation
+python trainer/train_tagger.py --filename tmp/TrigramTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction 0.99 --default="-Neznan-" ./pos
+date
+# BrillTagger without evaluation
+python trainer/train_tagger.py --filename tmp/BrillTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction 0.99 --default="-Neznan-" --brill ./pos
+date
+# ClassifierBasedPOSTagger - NaiveBayes without evaluation
+python trainer/train_tagger.py --filename tmp/NaiveBayes.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction 0.99 --default="-Neznan-" --sequential='' --classifier="NaiveBayes" ./pos
+date

### 2011-12-29 generateTaggers.sh

@@ -1,11 +1,8 @@
-# For evaluating the taggers after construction remove "--no-eval" argument.
-# By defult this evaluate on the whole corpus.
-# To evaluate tagger on only a fraction of a corpus (which is NOT used for training)
-# use argument "--fraction FRACTION"
-
 #
 # GENERATION
 #
+
+mkdir slovene_taggers
 
 date
 # TrigramTagger without evaluation
@@ -16,21 +13,4 @@ python trainer/train_tagger.py --filename slovene_taggers/BrillTagger.pickle --r
 date
 # ClassifierBasedPOSTagger - NaiveBayes without evaluation
 python trainer/train_tagger.py --filename slovene_taggers/NaiveBayes.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --default="-Neznan-" --sequential='' --classifier="NaiveBayes" --no-eval ./pos
-date
-
-#
-# EVALUATION (--fraction 0.99)
-# use 99% for generating the tagger and 1% for evaluation 
-# this is only used for evaluation, for real generation look above
-#
-
-date
-# TrigramTagger without evaluation
-python trainer/train_tagger.py --filename tmp/TrigramTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction 0.99 --default="-Neznan-" ./pos
-date
-# BrillTagger without evaluation
-python trainer/train_tagger.py --filename tmp/BrillTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction 0.99 --default="-Neznan-" --brill ./pos
-date
-# ClassifierBasedPOSTagger - NaiveBayes without evaluation
-python trainer/train_tagger.py --filename tmp/NaiveBayes.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction 0.99 --default="-Neznan-" --sequential='' --classifier="NaiveBayes" ./pos
 date

### 2011-12-29 evaluateTaggers.sh
Implemented loop for more fractions

@@ -8,16 +8,25 @@
 # use 99% for generating the tagger and 1% for evaluation 
 # this is only used for evaluation, for real generation look above
 #
+# for writing to console and file use:
+# ./evaluateTaggers.sh 2>&1 | tee evaluation.log
+#
 
 mkdir tmp
 
-date
-# TrigramTagger without evaluation
-python trainer/train_tagger.py --filename tmp/TrigramTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction 0.99 --default="-Neznan-" ./pos
-date
-# BrillTagger without evaluation
-python trainer/train_tagger.py --filename tmp/BrillTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction 0.99 --default="-Neznan-" --brill ./pos
-date
-# ClassifierBasedPOSTagger - NaiveBayes without evaluation
-python trainer/train_tagger.py --filename tmp/NaiveBayes.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction 0.99 --default="-Neznan-" --sequential='' --classifier="NaiveBayes" ./pos
-date
+for frac in 0.99 0.98 0.97 0.96 0.95
+do
+   echo "******************************************************************************************"
+   echo "*****************************    Fraction is $frac    *************************************"                              
+   echo "******************************************************************************************"
+   date
+   # TrigramTagger without evaluation
+   python trainer/train_tagger.py --filename tmp/TrigramTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction    $frac --default="-Neznan-" ./pos
+   date
+   # BrillTagger without evaluation
+   python trainer/train_tagger.py --filename tmp/BrillTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction $frac --default="-Neznan-" --brill ./pos
+   date
+   # ClassifierBasedPOSTagger - NaiveBayes without evaluation
+   python trainer/train_tagger.py --filename tmp/NaiveBayes.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction $frac --default="-Neznan-" --sequential='' --classifier="NaiveBayes" ./pos
+   date
+done

### 2011-12-29 generateTaggers.sh

@@ -1,5 +1,7 @@
 #
 # GENERATION
+# for writing to console and file use:
+# ./generateTaggers.sh 2>&1 | tee generation.log
 #
 
 mkdir slovene_taggers

### 2011-12-29 evaluateTaggers.sh
Modified evaluation

@@ -9,24 +9,34 @@
 # this is only used for evaluation, for real generation look above
 #
 # for writing to console and file use:
-# ./evaluateTaggers.sh 2>&1 | tee evaluation.log
+# ./evaluateTaggers.sh 2>&1 | tee evaluation.txt
 #
 
 mkdir tmp
 
 for frac in 0.99 0.98 0.97 0.96 0.95
 do
-   echo "******************************************************************************************"
-   echo "*****************************    Fraction is $frac    *************************************"                              
-   echo "******************************************************************************************"
+   echo "------------------------------------------------------------------------------------------------------------------"
+   echo "-----------------------------    Fraction is $frac    -------------------------------------------------------------"                              
+   echo "------------------------------------------------------------------------------------------------------------------"
+   echo -n "--- TIME: "
    date
+   echo "------------------------------------------------------------------------------------------------------------------"
    # TrigramTagger without evaluation
-   python trainer/train_tagger.py --filename tmp/TrigramTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction    $frac --default="-Neznan-" ./pos
+   python trainer/train_tagger.py --filename tmp/TrigramTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction  $frac --default="-Neznan-" ./pos
+   echo "------------------------------------------------------------------------------------------------------------------"
+   echo -n "--- TIME: "
    date
+   echo "------------------------------------------------------------------------------------------------------------------"
    # BrillTagger without evaluation
-   python trainer/train_tagger.py --filename tmp/BrillTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction $frac --default="-Neznan-" --brill ./pos
+   python trainer/train_tagger.py --filename tmp/BrillTagger.pickle --reader nltk.corpus.reader.tagged.TaggedCorpusReader --fileids '.+\.pos' --fraction $frac  --default="-Neznan-" --brill ./pos
+   echo "------------------------------------------------------------------------------------------------------------------"
+   echo -n "--- TIME: "
    date
+   echo "------------------------------------------------------------------------------------------------------------------"
[…]

### 2011-12-29 generateTaggers.sh

@@ -1,7 +1,7 @@
 #
 # GENERATION
 # for writing to console and file use:
-# ./generateTaggers.sh 2>&1 | tee generation.log
+# ./generateTaggers.sh 2>&1 | tee generation.txt
 #
 
 mkdir slovene_taggers

### 2011-12-29 paper/paper.tex
paper
Added chapter JOS corpora.

@@
 Our goal was to implement NLTK tagger for Slovene language.
-The resault of this project depends on two other larger projects.
+The result of this project depends on two other larger projects.
 First is \textit{JOS}\cite{JOS}(in Slovene it stands for: "Jezikoslovno Označevanje Slovenskega jezika"), from which we used the corpuses (large and structured set of texts).
-Second, \textit{nltk-trainer}\cite{nltk-trainer}, is an open-source project hosting on Github, which was used for actuall training of the tagger on Slovene sentences.
+Second, \textit{nltk-trainer}\cite{nltk-trainer}, is an open-source project hosting on Github, which was used for actual training of the tagger on Slovene sentences.
 Our work based mainly on putting all this pieces together into a working Slovene NLTK tagger.
-\section{JOS Corpus}
-\textit{TODO - BANIČ}
+\section{JOS corpora}
+The basic component for building Slovene tagger is \textbf{JOS corpora}.
+It contains collections of various text in Slovenian language with annotation.
+For our project we used jos1M corpus that contain 1 million words with it's appropriate lemmas and morphosyntactic descriptions.
+The JOS is compatible with Slovene\textit{ MULTEXT-East morphosyntactic specifications Version 4}\cite{MULTEXT-East}.
+It basic purpose is to define word classes.
+For each class there are various attributes and their appropriate values, which they can be mapped into morphosyntactic descriptions (i.e.
+MSD).
+The structure of JOS corpora is in  \textbf{Extensible Markup Language (XML)}.
+We can easily manipulate XML files in various programming languages (e.g.
+Python with xml.dom.minidom).
+This is mandatory because \textit{nltk-trainer}\cite{nltk-trainer} expects special form of input file.
+For further information see ~\ref{Corpus transformation}.
 \section{Nltk-trainer - NIKICC}
@@
 \subsection{Brill tagger}
-\textit{TODO - NIKICC}
+\textit{TODO - BANIČ}
 \subsection{Naive Bayes tagger}
@@
 \section{Evaluation}\label{Evaluation}
-\textit{TODO - NIKICC, BANIČ}
+\textit{TODO - NIKICC}
 \section{Conclusion}
@@
 \bibitem{nltk-trainer} \url{https://github.com/japerk/nltk-trainer}
-
+\bibitem{MULTEXT-East} \url{http://nl.ijs.si/ME/V4/msd/html/msd-sl.html}
 \end{thebibliography}

### 2011-12-29 evaluateTaggers.sh
moved evaluation resault into folder

@@ -9,7 +9,7 @@
 # this is only used for evaluation, for real generation look above
 #
 # for writing to console and file use:
-# ./evaluateTaggers.sh 2>&1 | tee evaluation.txt
+# ./evaluateTaggers.sh 2>&1 | tee evaluation/evaluation.txt
 #
 
 mkdir tmp

### 2011-12-31 evaluateTaggers.sh
Evaluation #2

@@ -14,7 +14,8 @@
 
 mkdir tmp
 
-for frac in 0.99 0.98 0.97 0.96 0.95
+#for frac in 0.99 0.98 0.97 0.96 0.95
+for frac in 0.94 0.93 0.92 0.91 0.90
 do
    echo "------------------------------------------------------------------------------------------------------------------"
    echo "-----------------------------    Fraction is $frac    -------------------------------------------------------------"                              

### 2012-01-01 evaluateTaggers.sh
Evalation 3

@@ -15,7 +15,8 @@
 mkdir tmp
 
 #for frac in 0.99 0.98 0.97 0.96 0.95
-for frac in 0.94 0.93 0.92 0.91 0.90
+#for frac in 0.94 0.93 0.92 0.91 0.90
+for frac in 0.88 0.86 0.84 0.82 0.80 0.75
 do
    echo "------------------------------------------------------------------------------------------------------------------"
    echo "-----------------------------    Fraction is $frac    -------------------------------------------------------------"                              

### 2012-01-05 paper/paper.tex
fixed referencing
PDF graph, reorganization of paper, IEEE style

@@
-\documentclass[11pt,a4paper,english,twocolumn]{article}
-%\usepackage[T1]{fontenc}
+\documentclass[11pt, conference, compsocconf]{IEEEtran}
+%\hyphenation{op-tical net-works semi-conduc-tor}
 \usepackage[utf8]{inputenc}
-\usepackage{babel}
-\usepackage{blindtext}
+\usepackage[pdftex]{graphicx} % za slike
 \usepackage{hyperref}
+\usepackage{multirow}
 
@@
 	\footnotesize University of Ljubljana \\
+	\footnotesize \texttt{niko.colneric@gmail.com} \\
 	\and
@@
 	\footnotesize University of Ljubljana \\
+	\footnotesize \texttt{MAIL@} \\
 }
@@
 \thispagestyle{empty}
-%\pagebreak
 
@@
 This paper describes the process of building Slovene tagger for use in NLTK.
+
+\section*{Keywords}
+
 \section{Introduction - NIKICC}
@@
 Our work based mainly on putting all this pieces together into a working Slovene NLTK tagger.
-\section{JOS corpora}
+
+\subsection{Trigram tagger}
+\textit{TODO - NIKICC}
+\subsection{Brill tagger}
+\textit{TODO - BANIČ}
+\subsection{Naive Bayes tagger}
+\textit{TODO - BANIČ}
+
+\section{Related work}
+\subsection{Nltk-trainer - NIKICC}
+Nltk-trainer\cite{nltk-trainer} is project, whose author is Jacob Perkins, from which we took the code for building the tagger.
+We almost exclusively used the script \textit{train\_tagger.py}, which has the ability to generate NLTK taggers.
+Its mandatory argument is corpus in .pos format.
+Other optional arguments can choose wheather to generate a Sequential Tagger, a Brill Tagger or a Classifier Based Tagger.
+It also has the possibility to set default tag (the tag for words, for which the tagger doesn't know how to tag) and wheather to evaluate the tagger or not.
+For evaluation of the tagger (section \ref{Evaluation}) this argument was used.
+
+\subsection{JOS corpora}
 The basic component for building Slovene tagger is \textbf{JOS corpora}.
@@
 This is mandatory because \textit{nltk-trainer}\cite{nltk-trainer} expects special form of input file.
-For further information see ~\ref{Corpus transformation}.
-\section{Nltk-trainer - NIKICC}
-Nltk-trainer\cite{nltk-trainer} is project, whose author is Jacob Perkins, from which we took the code for building the tagger.
-We almost exclusively used the script \textit{train\_tagger.py}, which has the ability to generate NLTK taggers.
-Its mandatory argument is corpus in .pos format.
-Other optional arguments can choose wheather to generate a Sequential Tagger, a Brill Tagger or a Classifier Based Tagger.
-It also has the possibility to set default tag (the tag for words, for which the tagger doesn't know how to tag) and wheather to evaluate the tagger or not.
-For evaluation of the tagger (section \ref{Evaluation}) this argument was used.
-\section{Taggers description}
-\subsection{Trigram tagger}
-\textit{TODO - NIKICC}
-\subsection{Brill tagger}
-\textit{TODO - BANIČ}
-\subsection{Naive Bayes tagger}
-\textit{TODO - BANIČ}
+For further information see \ref{Corpus transformation}.
+
 \section{Implementation}
-\subsection{Corpus transformation}
+\subsection{Corpus transformation}\label{Corpus transformation}
 \textit{TODO - BANIČ}
@@
 \textit{TODO - BANIČ}
-\section{Evaluation}\label{Evaluation}
+
+\section{Results}
 \textit{TODO - NIKICC}
+
+Average accuracy BrillTagger .
+Average accuracy NaiveBayes .
+Average accuracy TrigramTagger .
+
+\begin{table}[htb]
+\begin{center}
+\begin{tabular}{c|c|c|c}
+\textit{fraction} & \textit{Trigram} & \textit{NaiveBayes} & \textit{Brill} \\\hline\hline
+0.75 & 0.828648 & 0.845122 & 0.849031\\
+0.80 & 0.830353 & 0.846063 & 0.851387\\
+0.82 & 0.830551 & 0.846490 & 0.851527\\
+0.84 & 0.829855 & 0.846221 & 0.850839\\ 
+0.86 & 0.830609 & 0.847159 & 0.851240\\ 
+0.88 & 0.829901 & 0.847387 & 0.851680\\ 
+0.90 & 0.828807 & 0.846713 & 0.850737\\ 
+0.91 & 0.827759 & 0.846211 & 0.848899\\ 
+0.92 & 0.827613 & 0.846384 & 0.848964\\ 
+0.93 & 0.828121 & 0.846358 & 0.849090\\ 
+0.94 & 0.827826 & 0.846064 & 0.849636\\ 
+0.95 & 0.828913 & 0.847508 & 0.851041\\
+0.96 & 0.838798 & 0.853809 & 0.860899\\
+0.97 & 0.840060 & 0.856799 & 0.861213\\
+0.98 & 0.841683 & 0.858779 & 0.862869\\
+0.99 & 0.837875 & 0.856699 & 0.857844\\ \hline\hline
+\textit{average} & \textbf{0.831711} & \textbf{0.848985} & \textbf{0.852931}\\
+\end{tabular}
+\end{center}
+\caption{Accuracy of the taggers for different fractions. Fraction tells percentage of corpus used for training the tagger, the rest was used for evaluation.}
+\label{tab:evaluation}
+\end{table}
+
+\begin{figure}[htb]
+\begin{center}
+\leavevmode
+\includegraphics[width=0.5\textwidth]{../evaluation/graph.pdf} 
+\end{center}
+\caption{Evaluation results, with different fractions}
+\label{fig:evaluation}
+\end{figure}
+
 \section{Conclusion}
@@
 
-%\tableofcontents\newpage
-%\blinddocument
+\section*{Acknowledgements}
 
-%\thispagestyle{empty}
 \begin{thebibliography}{99}

### 2012-01-06 paper/paper.tex
Paper, time spent for tagging, measurements graphs

--- 2011-12-29 paper/paper.tex
@@
-\documentclass[11pt,a4paper,english,twocolumn]{article}
-%\usepackage[T1]{fontenc}
+\documentclass[10pt, conference, compsocconf]{IEEEtran}
+%\hyphenation{op-tical net-works semi-conduc-tor}
 \usepackage[utf8]{inputenc}
-\usepackage{babel}
-\usepackage{blindtext}
+\usepackage[pdftex]{graphicx} % for images
 \usepackage{hyperref}
 
-\title{\textbf{Slovene NLTK Tagger}}
+\title{Slovene NLTK Tagger}
 \author{
@@
 \thispagestyle{empty}
-%\pagebreak
 
-\section*{Abstract - NIKICC}
-This paper describes the process of building Slovene tagger for use in NLTK.  
-\section{Introduction - NIKICC}
-Tagger processes a sequence of words and attaches a part of speech tag to each word.
-We implemented tagger for Slovene language in NLTK.
-NLTK stands for Natural Language Toolkit and is a library for use in Python (currently for version below 3, we used 2.7).
-NLTK already has some taggers for English and some other languages.
-Our goal was to implement NLTK tagger for Slovene language.
-The result of this project depends on two other larger projects.
-First is \textit{JOS}\cite{JOS}(in Slovene it stands for: "Jezikoslovno Označevanje Slovenskega jezika"), from which we used the corpuses (large and structured set of texts).
-Second, \textit{nltk-trainer}\cite{nltk-trainer}, is an open-source project hosting on Github, which was used for actual training of the tagger on Slovene sentences.
-Our work based mainly on putting all this pieces together into a working Slovene NLTK tagger.
-\section{JOS corpora}
+\section*{Abstract} %NIKICC
+This paper describes the process of building NLTK tagger for custom language. NLTK stands for Natural Language Toolkit, a library for use in Python (version 2.7).
+Tagger is a object, which processes a sequence of words and attaches a part of speech tag to each word.
+Our language of implemetation was Slovene.
+We constructed sequential \textbf{Trigram tagger}, \textbf{Brill tagger} and classifier based \textbf{Naive Bayes tagger}, which differ in tagging accuracy and time consumed for tagging.
+All this differences are also detaily compared.
+\\\\
+\textit{Keywords:} Tagger, corpus, training
+
+\section{Introduction} %NIKICC
+Let us begin with the description of implemented taggers.
+
+%NIKICC
+\subsection[Trigram tagger]{Trigram tagger\footnote{Definition and images summerized from \cite{NLTKBOOK}}}
+To understant how \textit{trigram} works, we should first understand \textit{unigram}.
+Unigram taggers are based on a simple statistical algorithm: \textit{for each token, assign the tag that is most likely for that particular token}.
+In the case of tagging with unigram tagger, we only consider the current token, in isolation from any larger context. The best we can do is tag each word with its \textit{a priori} most likely tag.
+\par
+An n-gram tagger is a generalization of a unigram\footnote{Unigram could also be reffered to as 1-gram} tagger whose context is the current word together with the part-of-speech tags of the $n-1$ preceding tokens, as shown in figure \ref{fig:trigram}.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{tag-context.png} 
+\end{center}
+\caption{Tagger Context}
+\label{fig:trigram}
+\end{figure}
+
+The tag to be chosen, $t_{n}$, is circled, and the context is shaded in grey.
+In the example of an n-gram tagger shown in figure \ref{fig:trigram}, we have $n=3$; that is, we consider the tags of the two preceding words in addition to the current word.
+An n-gram tagger picks the tag that is most likely in the given context.
+\par
+Altrough we use the name trigram all throughtout this paper, we acctually refer to sequential tagger, which consists of
+\textit{affix}\footnote{The Affix tagger learns prefix and suffix patterns to determine the part of speech tag for word.}, \textit{unigram}, \textit{bigram} and \textit{trigram} taggers in this order respectively.
+First we build affix, which is used as ad \textit{backoff-tagger} for unigram.
+Unigram is afterwards used as a backoff-tagger for bigram, which is finally used for trigram.
+
+\subsection{Brill tagger} %BANIČ
+\textit{TODO - BANIČ}
+Banič glej ta link : nltk.googlecode.com/svn/trunk/doc/book/ch05.html
+poglavje 5.6
+\subsection{Naive Bayes tagger} %BANIČ
+\textit{TODO - BANIČ}
+
+\section{Related work}
+\subsection{Nltk-trainer} %NIKICC
+The code for building the tagger was taken form \textit{Nltk-trainer}\cite{nltk-trainer} project, whose author is Jacob Perkins.
+We almost exclusively used the script \textit{train\_tagger.py}, which has the ability to generate NLTK taggers.
+Its mandatory argument is a corpus in .pos format.
+Other optional arguments can choose wheather to generate a Sequential Tagger, a Brill Tagger or a Classifier Based Tagger.
+It also has the possibility to set default tag and wheather to evaluate the tagger or not.
+This argument was used for evaluating constructed taggers (see section \ref{results}).
+
+\subsection{JOS corpora} %BANIČ
 The basic component for building Slovene tagger is \textbf{JOS corpora}.
@@
 It basic purpose is to define word classes.
-For each class there are various attributes and their appropriate values, which they can be mapped into morphosyntactic descriptions (i.e.
-MSD).
+For each class there are various attributes and their appropriate values, which they can be mapped into morphosyntactic descriptions (i.e. MSD).
 The structure of JOS corpora is in  \textbf{Extensible Markup Language (XML)}.
-We can easily manipulate XML files in various programming languages (e.g.
-Python with xml.dom.minidom).
+We can easily manipulate XML files in various programming languages (e.g. Python with xml.dom.minidom).
 This is mandatory because \textit{nltk-trainer}\cite{nltk-trainer} expects special form of input file.
-For further information see ~\ref{Corpus transformation}.
-\section{Nltk-trainer - NIKICC}
-Nltk-trainer\cite{nltk-trainer} is project, whose author is Jacob Perkins, from which we took the code for building the tagger.
-We almost exclusively used the script \textit{train\_tagger.py}, which has the ability to generate NLTK taggers.
-Its mandatory argument is corpus in .pos format.
-Other optional arguments can choose wheather to generate a Sequential Tagger, a Brill Tagger or a Classifier Based Tagger.
-It also has the possibility to set default tag (the tag for words, for which the tagger doesn't know how to tag) and wheather to evaluate the tagger or not.
-For evaluation of the tagger (section \ref{Evaluation}) this argument was used.
-\section{Taggers description}
-\subsection{Trigram tagger}
+For further information see \ref{Corpus transformation}.
+
+\section{Implementation}
+\subsection{Corpus transformation} %BANIČ
+\label{Corpus transformation} 
+\textit{TODO - BANIČ}
+\subsection{Training the tagger} %NIKICC
 \textit{TODO - NIKICC}
-\subsection{Brill tagger}
+\subsection{Usage with NLTK} %BANIČ
 \textit{TODO - BANIČ}
-\subsection{Naive Bayes tagger}
-\textit{TODO - BANIČ}
-\section{Implementation}
-\subsection{Corpus transformation}
-\textit{TODO - BANIČ}
-\subsection{Training the tagger}
-\textit{TODO - NIKICC}
-\subsection{Usage with NLTK}
-\textit{TODO - BANIČ}
-\subsection{Encoding problems}
-\textit{TODO - BANIČ}
-\section{Evaluation}\label{Evaluation}
-\textit{TODO - NIKICC}
-\section{Conclusion}
+\subsection{Encoding problems} %BANIČ
 \textit{TODO - BANIČ}
 
-%\tableofcontents\newpage
-%\blinddocument
+\section{Results} %NIKICC
+\label{results}
+For all constructed taggers, we here provide their accuracy. First we divided the learning set into two groups (\textit{fraction} is the procentage of sentances assigned to first group).
+Now we construct the tagger on the first group and then evaluate its accuracy with the second group.
+The results, which were obtained by repeating this proces for various fractions, are shown in table \ref{tab:evaluation} and figure \ref{fig:evaluation}.
 
-%\thispagestyle{empty}
+\begin{table}[htb]
+\begin{center}
+\begin{tabular}{c|c|c|c}
+\textit{fraction} & \textit{Trigram} & \textit{NaiveBayes} & \textit{Brill} \\\hline\hline
+0.75  &  0.8286  &  0.8451  &  0.8490\\
+0.80  &  0.8304  &  0.8461  &  0.8514\\
+0.82  &  0.8306  &  0.8465  &  0.8515\\
+0.84  &  0.8299  &  0.8462  &  0.8508\\
+0.86  &  0.8306  &  0.8472  &  0.8512\\
+0.88  &  0.8299  &  0.8474  &  0.8517\\
+0.90  &  0.8288  &  0.8467  &  0.8507\\
+0.91  &  0.8278  &  0.8462  &  0.8489\\
+0.92  &  0.8276  &  0.8464  &  0.8490\\
+0.93  &  0.8281  &  0.8464  &  0.8491\\
+0.94  &  0.8278  &  0.8461  &  0.8496\\
+0.95  &  0.8289  &  0.8475  &  0.8510\\
+0.96  &  0.8388  &  0.8538  &  0.8609\\
+0.97  &  0.8401  &  0.8568  &  0.8612\\
+0.98  &  0.8417  &  0.8588  &  0.8629\\
+0.99  &  0.8379  &  0.8567  &  0.8578\\\hline\hline
+\textit{average} & \textbf{0.8317} & \textbf{0.8490} & \textbf{0.8529}
+\end{tabular}
+\end{center}
+\caption{Taggers accuracy for various fractions.}
+\label{tab:evaluation}
+\end{table}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{../evaluation/graph.pdf} 
+\end{center}
+\caption{Accuracy evaluation results from table \ref{tab:evaluation}.}
+\label{fig:evaluation_speed}
+\end{figure}
+
+During this evaluation we discover an other paramater, which should be taken into account.
+Under the assumption that most of the time was actually spent for tagging the sentances (besides just testing for equallity and percentage calculation are needed), we noticed that the time for evaluations differ drastically.
+The results of resarching this observation are listed in table \ref{tab:evaluation_speed} and figure \ref{fig:evaluation_speed}.
+The difference is enormous.
+Brill and Trigram run at almost the same speed, while Naive Bayes runs much slower.
+
+\begin{table}[htb]
+\begin{center}
+\begin{tabular}{c|c|c|c}
+\textit{number of words} & \textit{Trigram} & \textit{NaiveBayes} & \textit{Brill} \\\hline\hline
+   50  &  0.0003 &   4.1543  &  0.0011\\
+   75  &  0.0004 &   6.2464  &  0.0014\\
+  100  &  0.0006 &   8.5232  &  0.0018\\
+  125  &  0.0008 &  10.6370  &  0.0020\\
+  150  &  0.0009 &  12.4937  &  0.0024\\
+  175  &  0.0011 &  14.5484  &  0.0027\\
+  200  &  0.0011 &  16.3944  &  0.0029\\
+  225  &  0.0013 &  18.6332  &  0.0032\\
+  250  &  0.0015 &  20.7357  &  0.0035\\
+  275  &  0.0017 &  22.8280  &  0.0041\\
+  300  &  0.0018 &  24.8243  &  0.0042\\
+  325  &  0.0018 &  27.4270  &  0.0043\\
+  350  &  0.0022 &  29.2458  &  0.0051\\
+  400  &  0.0023 &  32.9599  &  0.0052\\
+  425  &  0.0024 &  36.0322  &  0.0056\\
+  450  &  0.0027 &  38.1471  &  0.0062\\
+  475  &  0.0027 &  39.2278  &  0.0064\\
+  500  &  0.0029 &  41.2688  &  0.0063\\
+\end{tabular}
+\end{center}
+\caption{Time spent in seconds for tagging different numbers of words.}
+\label{tab:evaluation_speed}
+\end{table}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{../evaluation/graph_speed.pdf} 
+\end{center}
+\caption{Time measurements from table \ref{tab:evaluation_speed}. NOTE: The green line is not visible because it is right under the red one.}
+\label{fig:evaluation}
+\end{figure}
+
+\section{Conclusion} %BANIČ
+\textit{TODO - BANIČ}
+
+\section*{Acknowledgements}
+
+\section{Currently not used}
+The result of this project depends on two other larger projects.
+First is \textit{JOS}\cite{JOS}(in Slovene it stands for: "Jezikoslovno Označevanje Slovenskega jezika"), from which we used the corpuses (large and structured set of texts).
+Second, \textit{nltk-trainer}\cite{nltk-trainer}, is an open-source project hosting on Github, which was used for actual training of the tagger on Slovene sentences.
+Our work based mainly on putting all this pieces together into a working Slovene NLTK tagger.
+
 \begin{thebibliography}{99}
@@
 \bibitem{MULTEXT-East} \url{http://nl.ijs.si/ME/V4/msd/html/msd-sl.html}
+\bibitem{NLTKBOOK} \url{http://www.nltk.org/book}
 \end{thebibliography}

### 2012-01-06 paper/paper.tex
More work on paper

@@
-\documentclass[10pt, conference, compsocconf]{IEEEtran}
-%\hyphenation{op-tical net-works semi-conduc-tor}
+\documentclass[11pt, conference, compsocconf]{IEEEtran}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{hyperref}
+\usepackage{listings}
 
@@
 
+\lstset{
+	language=bash,
+	basicstyle={\ttfamily\small},
+	xleftmargin=17pt,
+	breaklines=true,
+	extendedchars=\true,
+	frame=single
+}
+
 \begin{document}
@@
 All this differences are also detaily compared.
+We recommend using the Brill tagger.
 \\\\
-\textit{Keywords:} Tagger, corpus, training
+\textit{Keywords:} Tagger, corpus, training, tagging accuracy, time consumed
 
@@
 \subsection{Nltk-trainer} %NIKICC
-The code for building the tagger was taken form \textit{Nltk-trainer}\cite{nltk-trainer} project, whose author is Jacob Perkins.
-We almost exclusively used the script \textit{train\_tagger.py}, which has the ability to generate NLTK taggers.
-Its mandatory argument is a corpus in .pos format.
-Other optional arguments can choose wheather to generate a Sequential Tagger, a Brill Tagger or a Classifier Based Tagger.
-It also has the possibility to set default tag and wheather to evaluate the tagger or not.
-This argument was used for evaluating constructed taggers (see section \ref{results}).
+The code for building the tagger was taken from \textit{Nltk-trainer}  project \cite{nltk-trainer}, whose author is Jacob Perkins.
+Its an open-source project hosting on Github, with moto to \textit{Train NLTK objects with zero code}.
+We almost exclusively used the script \texttt{train\_tagger.py}, which has the ability to generate NLTK taggers. The
 
-\subsection{JOS corpora} %BANIČ
-The basic component for building Slovene tagger is \textbf{JOS corpora}.
+\subsection{JOS corpus} %BANIČ
+The basic component for building Slovene tagger is \textbf{JOS corpus} from JOS project \cite{JOS}.
+It stands for: \textit{Jezikoslovno Označevanje Slovenskega jezika} (Linguistic annotation of Slovenian language).
 It contains collections of various text in Slovenian language with annotation.
@@
 The structure of JOS corpora is in  \textbf{Extensible Markup Language (XML)}.
+
+\section{Implementation}
+\subsection{Corpus transformation} %BANIČ
+\label{Corpus transformation} 
+\textit{TODO - BANIČ}\\
 We can easily manipulate XML files in various programming languages (e.g. Python with xml.dom.minidom).
@@
 
-\section{Implementation}
-\subsection{Corpus transformation} %BANIČ
-\label{Corpus transformation} 
-\textit{TODO - BANIČ}
 \subsection{Training the tagger} %NIKICC
-\textit{TODO - NIKICC}
+As mentioned above, the script \texttt{train\_tagger.py} was used for this pupose.
+Its mandatory argument is a corpus in .pos format, which generation is described in \ref{Corpus transformation}.
+Beside this, there are many optional arguments.
+Here we will describe basics, for further description run script help with
+\begin{lstlisting}
+train_tagger.py --help
+\end{lstlisting}
+\par
+One is wheather to generate a Sequential Tagger, a Brill Tagger or a Classifier Based Tagger.
+For Sequential we can then select sequential backoff algorithm.
+This can be any combination of the following letters:
+\begin{itemize}
+\item a: AffixTagger
+\item u: UnigramTagger
+\item b: BigramTagger
+\item t: TrigramTagger
+\end{itemize}
+The default is \textit{aubt}, which was used for our Trigram tagger, but you can set this to the empty string to not train a sequential backoff tagger.
+The Brill Tagger is trained in front of the other tagger, which in our implementation was the Trigram tagger.
+When building classifier based algorithm, we must select the classifier or the sequence of classifier.
+Following are supported: NaiveBayes, DecisionTree, Maxent, GIS, IIS, CG, BFGS, Powell, LBFGSB, Nelder-Mead, MEGAM and TADM.
+We should mention, that the time spent for training classifier based tagger, is much larger than, the time for a sequential or brill.
+\par
+Script also has the possibility to set default tag. We used \textit{-Neznan-}, which is Slovene word for \textit{-unknown-}.
+\par
+Last argument, that should be mentioned, is wheather to evaluate the tagger or not and the procentage of sentaces to be used.
+The default value is 1, which means that the \textit{same} sentances are first used for training the tagger and then for evaluation.
+We reccomend this argument to be set below 1.
+If so, the right approach is used.
+Sentances are devided into two groups, one in used for training and second for evaluation. \textit{Fraction} is the procentage of sentances assigned to training group. 
+This was used for evaluating constructed taggers. For results of this evaluation see section \ref{results}).
+
 \subsection{Usage with NLTK} %BANIČ
@@
 \label{results}
-For all constructed taggers, we here provide their accuracy. First we divided the learning set into two groups (\textit{fraction} is the procentage of sentances assigned to first group).
-Now we construct the tagger on the first group and then evaluate its accuracy with the second group.
-The results, which were obtained by repeating this proces for various fractions, are shown in table \ref{tab:evaluation} and figure \ref{fig:evaluation}.
+For all constructed taggers, we here provide their accuracy.
+The results were obtained by repeating the proces of building the tagger with various evaluation arguments.
+See table \ref{tab:evaluation} and figure \ref{fig:evaluation}.
 
@@
 \end{center}
-\caption{Accuracy evaluation results from table \ref{tab:evaluation}.}
-\label{fig:evaluation_speed}
+\caption{Accuracy evaluation results.}
+\label{fig:evaluation}
 \end{figure}
-
-During this evaluation we discover an other paramater, which should be taken into account.
+\par
+To conclude, there are no major differences in tagging accuracy. The Brill and the Naive Bayes are about 2\% better than Trigram.
+\par
+However, during this evaluation we discover an other paramater, which should be taken into account.
 Under the assumption that most of the time was actually spent for tagging the sentances (besides just testing for equallity and percentage calculation are needed), we noticed that the time for evaluations differ drastically.
-The results of resarching this observation are listed in table \ref{tab:evaluation_speed} and figure \ref{fig:evaluation_speed}.
-The difference is enormous.
-Brill and Trigram run at almost the same speed, while Naive Bayes runs much slower.
+The results of resarching time consumption are listed in table \ref{tab:evaluation_speed} and graphically represented on figure \ref{fig:evaluation_speed}.
 
@@
 \end{center}
-\caption{Time measurements from table \ref{tab:evaluation_speed}. NOTE: The green line is not visible because it is right under the red one.}
-\label{fig:evaluation}
+\caption{Spent time measurements. NOTE: The green line is not visible because it is right under the red one.}
+\label{fig:evaluation_speed}
 \end{figure}
+
+The difference here is enormous. Brill and Trigram run at almost the same speed, while Naive Bayes runs extremely slower.
 
@@
 \section*{Acknowledgements}
-
-\section{Currently not used}
-The result of this project depends on two other larger projects.
-First is \textit{JOS}\cite{JOS}(in Slovene it stands for: "Jezikoslovno Označevanje Slovenskega jezika"), from which we used the corpuses (large and structured set of texts).
-Second, \textit{nltk-trainer}\cite{nltk-trainer}, is an open-source project hosting on Github, which was used for actual training of the tagger on Slovene sentences.
-Our work based mainly on putting all this pieces together into a working Slovene NLTK tagger.
+We would like to express our our gratitute to all the contributor of \textit{JOS} \cite{JOS} and  \textit{nltk-trainer} \cite{nltk-trainer} projects and out mentor at Faculty of Computer and Information Science.
 

### 2012-01-07 paper/paper.tex
Added my email to paper.
Added unfinished section Brill tagger.

@@
 	\footnotesize University of Ljubljana \\
-	\footnotesize \texttt{MAIL@} \\
+	\footnotesize \texttt{banic.nejc@gmail.com} \\
 }
@@
 poglavje 5.6
+The Brill tagger is another part-of-speech tagger and was first described by Eric Brill in 1993.
+Before trying to explain it, we must address issues with all n-gram taggers (e.g.
+\textit{Trigram tagger}).
+
+First issue with n-gram tagger is size of their n-gram table. If we wan't to use the tagger on mobile computing device, compromise must be taken between performance and model size.
+
+A second issue with n-gram tagger is that the only information it considers from prior context is tags.
+This is not practical because words themselves are useful information also.
+Therefore n-gram taggers are impractical models to be used to identify words in context.
+
+To bypass these problems we can use \textit{Brill tagger}.
+It can be summarised as an transformation-basede learning and performs very well using models that are just a fraction of the size of n-gram taggers.
+
+The general idea is very simple: guess the tag of each word, then go back and fix the mistakes.
+In this way, a Brill tagger successively transforms a bad tagging of a text into a better one.
+As with n-gram tagging, this is a supervised learning method, since we need annotated training data to figure out whether the tagger's guess is a mistake or not.
+However, unlike n-gram tagging, it does not count observations but compiles a list of transformational correction rules.
+
 \subsection{Naive Bayes tagger} %BANIČ

### 2012-01-08 paper/paper.tex
Added one additional section.

--- 2012-01-06 paper/paper.tex
@@
 \usepackage{listings}
-
+\usepackage[T1]{fontenc}
 \title{Slovene NLTK Tagger}
@@
 \subsection{Brill tagger} %BANIČ
-\textit{TODO - BANIČ}
-Banič glej ta link : nltk.googlecode.com/svn/trunk/doc/book/ch05.html
-poglavje 5.6
+
+The Brill tagger is another part-of-speech tagger and was first described by Eric Brill in 1993. Before trying to explain it, we must address issues with n-gram taggers (e.g. \textit{Trigram tagger}).
+
+First issue with n-gram taggers is size of their n-gram table. If we wan't to use tagging process on mobile computing device, compromise must be taken between performance and model size.
+
+A second issue with n-gram tagger is that the only information it considers from prior context is tags.
+This is not practical because words themselves are useful information also.
+Therefore n-gram taggers are impractical models to be used to identify words in context.
+
+To bypass those issues we can use \textit{Brill tagger}.
+It can be summarised as an transformation-based learning and performs very well using models that are just a fraction of the size of n-gram taggers.
+
+Despite obvious differences between Brill tagger and n-gram tagger there is one similarity.
+Both of them use \textit{supervised learning} method, since we require annotated training data to make sure whether the tagger's guess is a mistake or not.
+
+The idea behind the Brill tagger is: \textit{guess the tag of each word, then go back and fix the mistakes}. Basically, Brill tagger can transform bad tagging into a better one.
+
 \subsection{Naive Bayes tagger} %BANIČ
@@
 \subsection{JOS corpus} %BANIČ
-The basic component for building Slovene tagger is \textbf{JOS corpus} from JOS project \cite{JOS}.
+The basic component for building Slovene tagger is \textit{JOS corpus} from JOS project \cite{JOS}.
 It stands for: \textit{Jezikoslovno Označevanje Slovenskega jezika} (Linguistic annotation of Slovenian language).
@@
 This is mandatory because \textit{nltk-trainer}\cite{nltk-trainer} expects special form of input file.
-For further information see \ref{Corpus transformation}.
+
+The first step for corpus transformation is to identify proper \textit{XML tags}\cite{xml_tags} in JOS corpora.
+There are several types, but we must parse only those in \textit{sentence tag }(\textit{<s>}):
+\begin{itemize}
+\item[•]\textit{<S>} - space.
+\item[•]\textit{<c>} - character i.e. punctuation marks (e.g. period (.), comma (,), colon (:), ... ).
+\item[•]\textit{<w>} - word.
+\item[•]\textit{<term>} - a word or group of words designating something, especially in a particular field. 
+\end{itemize}
+
+Trainer's mandatory argument is a corpus in .pos format. It contains words with their morphosyntactic descriptions and punctuation marks. The correct syntax is shown in the next example:
+\texttt{Ameriški/Ppnmeid CNN/Slmei ,/,  brez/Dr }
+
+The trainer will train the .pos properly if:
+\begin{itemize}
+\item[•] The words and their MSD are separated with slash.
+\item[•] Every punctuation mark is separated with slash and then follows the same punctuation mark.
+\item[•] Words and punctuation marks are between space.
+\end{itemize}
+
+If we follow these "rules", implementation is simple.
+We just need to go through all child tags in sentence tag and properly write them in a file.
+Because Slovenian language contain letters with caron, proper encoding of tags must be provided.
+This is simply done with function \textit{encode()} (e.g.
+encode("utf-8") - this is proper encoding for Slovenian language).
+\textit{Jos1M} corpus is build up with 10 XML files (e.g. \textit{jos1M-01}, \textit{jos1M-02}, ...), so we must parse all of them. 
+File \texttt{transformJOS.py} contains the corpus transformation.
 
 \subsection{Training the tagger} %NIKICC
-As mentioned above, the script \texttt{train\_tagger.py} was used for this pupose.
+As mentioned above, the script \texttt{train\_tagger.py} was used for this purpose.
 Its mandatory argument is a corpus in .pos format, which generation is described in \ref{Corpus transformation}.
@@
 The default value is 1, which means that the \textit{same} sentances are first used for training the tagger and then for evaluation.
-We reccomend this argument to be set below 1.
+We recommend this argument to be set below 1.
 If so, the right approach is used.
@@
 \textit{TODO - BANIČ}
+We can use the trained tagger and NLTK to see how they can be used in Slovenian language.
+The requirements are installed NLTK and Python 2.7.
+We set an example in file \texttt{example.py}.
+First we need to import nltk and \textit{pickle}, which is module for serialization and de-serialization a Python object structure ("pickling" - Python object hierarchy is converted into a byte stream).
+Pickle is used for loading classifiers such as Trigram tagger, Brill tagger, etc.
+
+We also implemented a function that returns two description dictionaries for  MSDs with attribute=value expansions in Slovene and English. That way users can view which part-of-speech words belong to.
 \subsection{Encoding problems} %BANIČ
@@
 \bibitem{NLTKBOOK} \url{http://www.nltk.org/book}
+\bibitem{xml_tags} \url{http://nl.ijs.si/ssj/schema/tei_ssj_doc.pdf}
 \end{thebibliography}

### 2012-01-08 paper/paper.tex
Completed tagger and fixed example.py.

@@
-\documentclass[11pt, conference, compsocconf]{IEEEtran}
+\documentclass[10pt, conference, compsocconf]{IEEEtran}
 \usepackage[utf8]{inputenc}
@@
 Our language of implemetation was Slovene.
-We constructed sequential \textbf{Trigram tagger}, \textbf{Brill tagger} and classifier based \textbf{Naive Bayes tagger}, which differ in tagging accuracy and time consumed for tagging.
+We constructed sequential \textit{Trigram tagger}, \textit{Brill tagger} and classifier based \textit{Naive Bayes tagger}, which differ in tagging accuracy and time consumed for tagging.
 All this differences are also detaily compared.
@@
 
-\subsection{Naive Bayes tagger} %BANIČ
-\textit{TODO - BANIČ}
+\subsection[Naive Bayes classifier]{Naive Bayes classifier\footnote{Definition and images summerized from \cite{NLTKBOOK}}}
+Both Brill and Trigram are sequential backoff taggers. Meanwhile, Naive Bayes is \textit{classifier based tagger}.
+Classification is a process where, for a given output, we have to choose the correct class. Basically, it is an  procedure for assigning  input data into one of a given number of categories.
+
+Examples of classification tasks are:
+\begin{itemize}
+\item[•] whether an email is spam or not.
+\item[•] diagnosis to a given patient as described by observed characteristics of the patient
+\end{itemize}
+The basic classification task has a number of interesting variants.
+For example, in multi-class classification, each instance may be assigned multiple labels; in open-class classification, the set of labels is not defined in advance; and in sequence classification, a list of inputs are jointly classified.
+
+Every feature in naive Bayes classifier can determine which label can be assigned to a given input.
+Choosing a label for an input is an process, that consists of calculating the prior probability of each label (it is determined by calculating frequency in training set).
+Priori probability is combined with contribution from each feature.
+This is necessary to get the likelihood estimate for each label.
+Input value is then assigned the label whose likelihood estimate is the highest.
+The process is shown in \ref{fig:naive-bayes-triangle}.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{naive-bayes-triangle.png} 
+\end{center}
+\caption{An illustration of the procedure used by the naive Bayes classifier to choose the topic for a document.}
+\label{fig:naive-bayes-triangle}
+\end{figure}
+
+Let's assume that in the training corpus most documents have automotive topic.
+Because of this assumption, the classifier starts near "automotive" label.
+Later it must take into consideration the effect of each feature.
+Example considers, that input document contains word "dark", which indicates for "murder mystery", but also contains the word "football", which (strongly) indicate for "sports".
+After every feature has made its contribution, the classifier checks which label it is closest to, and assigns that label to the input.
+
+Features make their contribution to the  decision by voting against those labels, that don't occur with this feature.
+Likelihood score is therefore reduced for each label by multiplying it with the probability that an input value with that label would have that feature.
+Suppose that the word occurs ran occurs in  12\% of the sports documents, 10\% of the murder mystery documents, and 2\% of the automotive documents.
+The likelihood score for the sports label will be multiplied by 0.12, the likelihood score for the murder mystery label will be multiplied by 0.1, and the likelihood score for the automotive label will be multiplied by 0.02.
+The overall effect will be to slightly reduce the score of the sports label, reduce the murder mystery label (more then sports label) and significantly reduce the automotive label.
+Figure \ref{fig:naive-bayes-bargraph} illustrates this example.
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=0.4\textwidth]{naive_bayes_bargraph.png} 
+\end{center}
+\caption{Illustrating the process of calculating label likelihoods with naive Bayes.}
+\label{fig:naive-bayes-bargraph}
+\end{figure}
 
@@
 For each class there are various attributes and their appropriate values, which they can be mapped into morphosyntactic descriptions (i.e. MSD).
-The structure of JOS corpora is in  \textbf{Extensible Markup Language (XML)}.
+The structure of JOS corpora is in  \textit{Extensible Markup Language (XML)}.
 
@@
 \label{Corpus transformation}
-\textit{TODO - BANIČ}\\
+
 We can easily manipulate XML files in various programming languages (e.g. Python with xml.dom.minidom).
@@
 Trainer's mandatory argument is a corpus in .pos format. It contains words with their morphosyntactic descriptions and punctuation marks. The correct syntax is shown in the next example:
-\texttt{Ameriški/Ppnmeid CNN/Slmei ,/,  brez/Dr }
+\textit{Ameriški/Ppnmeid CNN/Slmei ,/,  brez/Dr }
 
@@
 \subsection{Usage with NLTK} %BANIČ
-\textit{TODO - BANIČ}
-We can use the trained tagger and NLTK to see how they can be used in Slovenian language.
-The requirements are installed NLTK and Python 2.7.
-We set an example in file \texttt{example.py}.
-First we need to import nltk and \textit{pickle}, which is module for serialization and de-serialization a Python object structure ("pickling" - Python object hierarchy is converted into a byte stream).
+
+We can use the trained tagger and NLTK to see how they can be used in Slovenian language. The requirements are installed NLTK and Python 2.7. We set an example in file \texttt{example.py}.
+First we need to import following functions:
+\begin{itemize}
+\item[•] \textit{pickle} - it is a module for serialization and de-serialization a Python object structure ("pickling" - Python object hierarchy is converted into a byte stream).
 Pickle is used for loading classifiers such as Trigram tagger, Brill tagger, etc.
 
-We also implemented a function that returns two description dictionaries for  MSDs with attribute=value expansions in Slovene and English. That way users can view which part-of-speech words belong to.
+\item[•] \textit{PunktWordTokenizer} \cite{tokenizer} - it  is just one of many NLTK's tokenizers, that can be used to divide strings into lists of substring.
+Tokenizers are subclasses of \textit{nltk.tokenizer.TokenizerI} interface, which defines the \textit{tokenizer()} method.
+PunktWordTokenizer uses regular expressions to divide a word into tokens.
+These tokens are then used by the tagger.
+Following example will demonstrate, how does PunktWordTokenizer tokenize a sentence.
+If we have the following sentence:
+
+\textit{Potem se je obrnil proti nam in skozi solze izoblikoval najlepši nasmeh, kar sem jih videl v življenju.}\\
+the tokenizer will output the following list of tokens:
+
+[\textit{'Potem', 'se', 'je', 'obrnil', 'proti', 'nam', 'in', 'skozi', 'solze', 'izoblikoval', 'najlepši', 'nasmeh', ',', 'kar', 'sem', 'jih', 'videl', 'v', 'življenju.'}]
+\item[•] \textit{nltk.data} - it is a module that contains various functions to load NLTK files, such as corpora.
+We use it for \textit{sentence segmentation}.
+In some cases user can input not a sentence, but a whole paragraph.
+Before using PunktWordTokenizer to tokenize sentence into tokens, we need to segment it into sentences.
+This can be easily used by including Slovene Punkt sentence segmenter.
+Following example shows, how does sentence segmenter work.
+If we have the following paragraph:
+
+\textit{Zdravljíca je pesem, ki jo je France Prešeren napisal novembra leta 1844.
+Po 6.
+členu ustave je Zdravljica besedilo himne Republike Slovenije.
+Melodija je iz istoimenske zborovske skladbe skladatelja Stanka Premrla.}\\
+
+the sentence segmenter will output the following list of sentences:
+
+[\textit{'Zdravljíca je pesem, ki jo je France Prešeren napisal novembra leta 1844.', 'Po 6.
+členu ustave je Zdravljica besedilo himne Republike Slovenije.', 'Melodija je iz istoimenske zborovske skladbe skladatelja Stanka Premrla.'}]
+\end{itemize}
+
+The first step is to use sentence segmenter, to form a list of sentences.
+This list is then being tokenized with PunktWordTokenizer.
+One drawback with this tokenizer is that it doesn't tokenize the last word of a sentence and the following period (as shown in the example).
+Before using the PunktWordTokenizer, we must divide them with space.
+After we tokenize the sentences, tagging can be used.
+The result is a list of tuples with words or characters and their appropriate morphosyntactic description.
+
+We also implemented a function that returns two description dictionaries for MSDs with attribute=value expansions in Slovene and English.
+This way user can view which part-of-speech words belong to in Slovene or English.
+Following example demonstrates the expected output:
+\\
+\textit{( Lep  |  PPNMEIN ) - pridevnik \\
+	( je  |  GP-STE-N ) - glagol\\
+	( dan  |  SOMEI ) - samostalnik\\
+	( ,  |  , ) - ni razlage\\
+	( vse  |  ZC-SEI ) - zaimek\\
+	( diši  |  GGNSTE ) - glagol\\
+	( že  |  L ) - členek\\
+	( po  |  DM ) - predlog\\
+	( pomladi  |  SOZEM ) - samostalnik\\
+	( !  |  ! ) - ni razlage}\\
+
+But there is an issue.
+Because Slovenian language contain letters with caron, the output will have unicode symbols for them.
+In the previous examples, the output is already corrected.
+How we solve this is described in the following section.
 \subsection{Encoding problems} %BANIČ
-\textit{TODO - BANIČ}
 
+Printing the result of our example is not optimal.
+Because we want users to know what these characters are, we implemented a solution.
+To resolve this, we must decode every word to \textit{UTF - 8} (because words we tag are in UTF - 8).
+If a user wants to print the result in "pretty" form, he should call function \textit{prettyPrint()} or \textit{prettyPrintByLine()} (both are implemented in \texttt{example.py}).
 \section{Results} %NIKICC
@@
 \par
-However, during this evaluation we discover an other paramater, which should be taken into account.
-Under the assumption that most of the time was actually spent for tagging the sentances (besides just testing for equallity and percentage calculation are needed), we noticed that the time for evaluations differ drastically.
+However, during this evaluation we discover an other parameter, which should be taken into account.
+Under the assumption that most of the time was actually spent for tagging the sentences (besides just testing for equality and percentage calculation are needed), we noticed that the time for evaluations differ drastically.
 The results of resarching time consumption are listed in table \ref{tab:evaluation_speed} and graphically represented on figure \ref{fig:evaluation_speed}.
@@
 \section{Conclusion} %BANIČ
-\textit{TODO - BANIČ}
+
+To conclude our paper, tagging in Slovene language is practical with proper use of Natural Language Toolkit and \textit{nltk-trainer}.
+Time spent in seconds for tagging different numbers of words, show that Brill and Trigram ran at the same speed, while Naive Bayes ran slow.
+There are no major differences in tagging accuracy (Brill and the Naive Bayes are about 2\% better than Trigram).
 
 \section*{Acknowledgements}
-We would like to express our our gratitute to all the contributor of \textit{JOS} \cite{JOS} and  \textit{nltk-trainer} \cite{nltk-trainer} projects and out mentor at Faculty of Computer and Information Science.
-
+We would like to express our gratitude to all the contributors of \textit{JOS} \cite{JOS} and  \textit{nltk-trainer} \cite{nltk-trainer} projects and our mentor at Faculty of Computer and Information Science.
+Also this project would not be possible without \textit{Natural Language Toolkit} \cite{nltk} and their admirable documentation.
+\\
 \begin{thebibliography}{99}
+\bibitem{nltk} \url{http://www.nltk.org/}
 \bibitem{JOS} \url{http://nl.ijs.si/jos/}
@@
 \bibitem{xml_tags} \url{http://nl.ijs.si/ssj/schema/tei_ssj_doc.pdf}
+\bibitem{tokenizer} \url{http://docs.huihoo.com/nltk/0.9.5/guides/tokenize.html}
 \end{thebibliography}

### 2012-01-08 example.py

@@ -1,4 +1,4 @@
-# -*- coding: utf-8 -*
+
 import pickle
 import nltk.data
 import time, datetime
@@ -7,18 +7,18 @@ from nltk import PunktWordTokenizer
 ### FUNCTIONS ###
 #prettPrint
 def prettyPrint(uniResult):
-    print ", ".join("(%s | %s)" % a for a in uniResult)
+    print ", ".join("(%s | %s)" % (x.decode("utf-8"),y) for (x,y) in uniResult)
 
 def prettyPrintByLine(uniResult):
-    for a in uniResult:
-        print "(%s | %s)" % a
+    for (x,y) in uniResult:
+        print "(",x.decode("utf-8")," | ",y,")"
 
 def prettyPrintWithDescription(uniResult, description):
     for (x,y) in uniResult:
 	if y in description:
-	    print "(",x," | ",y,") - "+unicode(description[y][0],"utf-8")
+	    print "(",x.decode("utf-8")," | ",y,") - "+unicode(description[y][0],"utf-8")
         else:
-	    print "(",x," | ",y,") - ni razlage"
+	    print "(",x.decode("utf-8")," | ",y,") - ni razlage"
 
 #slo tag - input UNICODE
 def sloTag(uText):
@@ -71,13 +71,14 @@ _t1 = time.time()
 
 
 #uText is input text
-#uText = u'Lep je dan, vse diši že po pomladi!'
-#uText = u'Živé naj vsi naródi, ki hrepené dočakat dan, da, koder sonce hodi, prepir iz svéta bo pregnan, ko rojak prost bo vsak, ne vrag, le sosed bo mejak.'
-uText = u'Potem se je obrnil proti nam in skozi solze izoblikoval najlepši nasmeh, kar sem jih videl v življenju.'
+#uText = 'Lep je dan, vse diši že po pomladi!'
+#uText = 'Živé naj vsi naródi, ki hrepené dočakat dan, da, koder sonce hodi, prepir iz svéta bo pregnan, ko rojak prost bo vsak, ne vrag, le sosed bo mejak.'
+uText = 'Potem se je obrnil proti nam in skozi solze izoblikoval najlepši nasmeh, kar sem jih videl v življenju.'
 
 
 _t2 = time.time()
 result = sloTag2(uText)
+print result
 _t3 = time.time()
 
 prettyPrintWithDescription(result, dict_sl)

### 2012-01-08 paper/paper.tex
Work on the paper

@@
 \section*{Abstract} %NIKICC
-This paper describes the process of building NLTK tagger for custom language. NLTK stands for Natural Language Toolkit, a library for use in Python (version 2.7).
-Tagger is a object, which processes a sequence of words and attaches a part of speech tag to each word.
+This paper describes the process of building NLTK tagger for a custom language. NLTK stands for Natural Language Toolkit, a library for use in Python (version 2.7).
+Tagger is an object, which processes a sequence of words and attaches a part of speech tag to each word.
 Our language of implemetation was Slovene.
@@
 \section{Introduction} %NIKICC
-Let us begin with the description of implemented taggers.
+First, let us say a few words about the structure of the paper.
+In this chapter, we provide a breif theorethical description of imlemented taggers.
+Next we will describe the related work in section \ref{relatedWork}, which was the base of our project.
+In section \ref{implementation} the proces of implementaion itself is explained.
+The example, how to use the tagger is provided, along with some discussion on encoding problems.
+Finally in section \ref{results} is the evaluation of accuracy and speed.
+\par
+Now we begin with the description of implemented taggers.
 
@@
 
-\begin{figure}[htb]
+\begin{figure}[h]
 \begin{center}
@@
 
-First issue with n-gram taggers is size of their n-gram table. If we wan't to use tagging process on mobile computing device, compromise must be taken between performance and model size.
+First issue with n-gram taggers is size of their n-gram table. If we want to use tagging process on mobile computing device, compromise must be taken between performance and model size.
 
-A second issue with n-gram tagger is that the only information it considers from prior context is tags.
+A second issue with n-gram tagger is, that the only information it considers from prior context is tags.
 This is not practical because words themselves are useful information also.
@@
 \begin{itemize}
-\item[•] whether an email is spam or not.
-\item[•] diagnosis to a given patient as described by observed characteristics of the patient
+\item whether an email is spam or not.
+\item diagnosis to a given patient as described by observed characteristics of the patient
 \end{itemize}
@@
 Every feature in naive Bayes classifier can determine which label can be assigned to a given input.
-Choosing a label for an input is an process, that consists of calculating the prior probability of each label (it is determined by calculating frequency in training set).
+Choosing a label for an input is a process, that consists of calculating the prior probability of each label (it is determined by calculating frequency in training set).
 Priori probability is combined with contribution from each feature.
@@
 \section{Related work}
+\label{relatedWork}
 \subsection{Nltk-trainer} %NIKICC
@@
 Its an open-source project hosting on Github, with moto to \textit{Train NLTK objects with zero code}.
-We almost exclusively used the script \texttt{train\_tagger.py}, which has the ability to generate NLTK taggers. The
+We almost exclusively used the script \texttt{train\_tagger.py}, which has the ability to generate NLTK taggers. This project made it a lot easier for us to train the taggers on Slovenian corpuses.
 
@@
 \section{Implementation}
+\label{implementation}
 \subsection{Corpus transformation} %BANIČ
@@
 We can easily manipulate XML files in various programming languages (e.g. Python with xml.dom.minidom).
-This is mandatory because \textit{nltk-trainer}\cite{nltk-trainer} expects special form of input file.
+This is mandatory because \textit{nltk-trainer}\cite{nltk-trainer} expects special form of input file (.pos file).
 
@@
 \begin{itemize}
-\item[•]\textit{<S>} - space.
-\item[•]\textit{<c>} - character i.e. punctuation marks (e.g. period (.), comma (,), colon (:), ... ).
-\item[•]\textit{<w>} - word.
-\item[•]\textit{<term>} - a word or group of words designating something, especially in a particular field. 
+\item \textit{<S>} - space.
+\item \textit{<c>} - character i.e. punctuation marks (e.g. period (.), comma (,), colon (:), ... ).
+\item \textit{<w>} - word.
+\item \textit{<term>} - a word or group of words designating something, especially in a particular field. 
 \end{itemize}
 
-Trainer's mandatory argument is a corpus in .pos format. It contains words with their morphosyntactic descriptions and punctuation marks. The correct syntax is shown in the next example:
-\textit{Ameriški/Ppnmeid CNN/Slmei ,/,  brez/Dr }
+For nltk-trainer to work, we must transform \texttt{jos1m.xml} to \texttt{.pos} file, which contains words with their morphosyntactic descriptions and punctuation marks.
+The correct syntax is shown in the next example:
+\begin{lstlisting}
+Britanski/Ppnmeid BBC/Slmei ,/,  brez/Dr 
+\end{lstlisting}
 
-The trainer will train the .pos properly if:
+The trainer will train the tagger properly, if in .pos file:
 \begin{itemize}
-\item[•] The words and their MSD are separated with slash.
-\item[•] Every punctuation mark is separated with slash and then follows the same punctuation mark.
-\item[•] Words and punctuation marks are between space.
+\item The words and their MSD are separated with slash.
+\item Every punctuation mark is separated with slash and then follows the same punctuation mark.
+\item Words and punctuation marks are between space.
 \end{itemize}
@@
 encode("utf-8") - this is proper encoding for Slovenian language).
-\textit{Jos1M} corpus is build up with 10 XML files (e.g. \textit{jos1M-01}, \textit{jos1M-02}, ...), so we must parse all of them. 
+\textit{Jos1M} corpus is build up with 10 XML files (e.g. \texttt{jos1M-01.xml},  \texttt{jos1M-02.xml}, ..., \texttt{jos1M-10.xml}), so we must parse all of them. 
 File \texttt{transformJOS.py} contains the corpus transformation.
@@
 Pickle is used for loading classifiers such as Trigram tagger, Brill tagger, etc.
+\begin{lstlisting}
+tagger = pickle.load( open("slovene_taggers/BrillTagger.pickle") )
+\end{lstlisting}
 
@@
 [\textit{'Potem', 'se', 'je', 'obrnil', 'proti', 'nam', 'in', 'skozi', 'solze', 'izoblikoval', 'najlepši', 'nasmeh', ',', 'kar', 'sem', 'jih', 'videl', 'v', 'življenju.'}]
+
+As you can see, there is the problem with the last dot. The solution we used is described below.
+
 \item[•] \textit{nltk.data} - it is a module that contains various functions to load NLTK files, such as corpora.
@@
 Following example demonstrates the expected output:
-\\
+\\\\
 \textit{( Lep  |  PPNMEIN ) - pridevnik \\
@@
 
+We must exmplain, that here only the part of whole MSD is printed. For example the whole MSD for \textit{Lep} is:\\\\
+\textit{pridevnik vrsta=splošni spol=moški število=ednina sklon=imenovalnik živost=0 vid=0 oblika=0 oseba=0 nikalnost=0 stopnja=nedoločeno določnost=ne število\_svojine=0 spol\_svojine=0 naslonskost=0 zapis=0}\\
+\par
+The first output was creted with function \textit{prettyPrintWithDescription}, full output can be created with \textit{prettyPrintWithFullDescription}. Both function are defined in \texttt{example.py}
+\par
 But there is an issue.
@@
 In the previous examples, the output is already corrected.
-How we solve this is described in the following section.
+Solution is described in the following section.
 \subsection{Encoding problems} %BANIČ
@@
 If a user wants to print the result in "pretty" form, he should call function \textit{prettyPrint()} or \textit{prettyPrintByLine()} (both are implemented in \texttt{example.py}).
+
 \section{Results} %NIKICC
@@
 For all constructed taggers, we here provide their accuracy.
-The results were obtained by repeating the proces of building the tagger with various evaluation arguments.
+The results were obtained by repeating the proces of building the tagger with various evaluation arguments (see script \texttt{evaluateTaggers.sh}).
 See table \ref{tab:evaluation} and figure \ref{fig:evaluation}.
 
-\begin{table}[htb]
+\begin{table}[h]
 \begin{center}
@@
 Under the assumption that most of the time was actually spent for tagging the sentences (besides just testing for equality and percentage calculation are needed), we noticed that the time for evaluations differ drastically.
-The results of resarching time consumption are listed in table \ref{tab:evaluation_speed} and graphically represented on figure \ref{fig:evaluation_speed}.
+The results of resarching time consumption are listed in table \ref{tab:evaluation_speed} and graphically represented on figure \ref{fig:evaluation_speed} (see script \texttt{evaluateTaggersSpeed.py}).
 
-\begin{table}[htb]
+\begin{table}[h]
 \begin{center}
@@
 
-The difference here is enormous. Brill and Trigram run at almost the same speed, while Naive Bayes runs extremely slower.
+The difference here is enormous. Brill and Trigram run at almost the same speed, while Naive Bayes runs extremely slower. This is also one of the reason why we prefer usig the Brill tagger.
 
 \section{Conclusion} %BANIČ
-
 To conclude our paper, tagging in Slovene language is practical with proper use of Natural Language Toolkit and \textit{nltk-trainer}.
-Time spent in seconds for tagging different numbers of words, show that Brill and Trigram ran at the same speed, while Naive Bayes ran slow.
-There are no major differences in tagging accuracy (Brill and the Naive Bayes are about 2\% better than Trigram).
+Time spent in seconds for tagging different numbers of words, show that Brill and Trigram ran at the same speed, while Naive Bayes ran drasticly slower.
+There are no major differences in tagging accuracy.
+However Brill and the Naive Bayes are about 2\% better than Trigram.
+\par
+Besides these three taggers we built, others can be built without any bigger modification.
+All that is needed is to set some arguments when calling the script \texttt{train\_tagger.py}.
+The main reason we built only these three is computing power.
+These three are the only one we could built and evaluate in some resonable time.
+\par
+We also contacted the NLTK developers group and will arrange our taggers to become the part of NLTK.
 
 \section*{Acknowledgements}
-We would like to express our gratitude to all the contributors of \textit{JOS} \cite{JOS} and  \textit{nltk-trainer} \cite{nltk-trainer} projects and our mentor at Faculty of Computer and Information Science.
+We would like to express our gratitude to all the contributors of \textit{JOS} \cite{JOS} project, \textit{nltk-trainer} \cite{nltk-trainer} project and to Marko Robnik-Šikonja, PhD, Associate Professor, who was our mentor at Faculty of Computer and Information Science.
 Also this project would not be possible without \textit{Natural Language Toolkit} \cite{nltk} and their admirable documentation.
-\\
+
 \begin{thebibliography}{99}

### 2012-01-08 example.py

@@ -1,3 +1,4 @@
+# -*- coding: utf-8 -*
 
 import pickle
 import nltk.data
@@ -17,6 +18,13 @@ def prettyPrintWithDescription(uniResult, description):
     for (x,y) in uniResult:
 	if y in description:
 	    print "(",x.decode("utf-8")," | ",y,") - "+unicode(description[y][0],"utf-8")
+        else:
+	    print "(",x.decode("utf-8")," | ",y,") - ni razlage"
+
+def prettyPrintWithFullDescription(uniResult, description):
+    for (x,y) in uniResult:
+	if y in description:
+	    print "(",x.decode("utf-8")," | ",y,") - "+unicode(description[y][0],"utf-8")+" "+unicode(description[y][1],"utf-8")
         else:
 	    print "(",x.decode("utf-8")," | ",y,") - ni razlage"
 
@@ -71,9 +79,9 @@ _t1 = time.time()
 
 
 #uText is input text
-#uText = 'Lep je dan, vse diši že po pomladi!'
+uText = 'Lep je dan, vse diši že po pomladi!'
 #uText = 'Živé naj vsi naródi, ki hrepené dočakat dan, da, koder sonce hodi, prepir iz svéta bo pregnan, ko rojak prost bo vsak, ne vrag, le sosed bo mejak.'
-uText = 'Potem se je obrnil proti nam in skozi solze izoblikoval najlepši nasmeh, kar sem jih videl v življenju.'
+#uText = 'Potem se je obrnil proti nam in skozi solze izoblikoval najlepši nasmeh, kar sem jih videl v življenju.'
 
 
 _t2 = time.time()

### 2012-01-08 paper/paper.tex
Whole MSD in table

@@
 
-We must exmplain, that here only the part of whole MSD is printed. For example the whole MSD for \textit{Lep} is:\\\\
-\textit{pridevnik vrsta=splošni spol=moški število=ednina sklon=imenovalnik živost=0 vid=0 oblika=0 oseba=0 nikalnost=0 stopnja=nedoločeno določnost=ne število\_svojine=0 spol\_svojine=0 naslonskost=0 zapis=0}\\
-\par
+We must exmplain, that here only the part of whole MSD is printed. For example the whole MSD for \textit{Lep} shown in table \ref{tab:MSDLep}.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{c|c}
+\multicolumn{2}{c}{pridevnik}\\\hline\hline
+vrsta & splošni \\
+spol & moški \\
+število & ednina \\
+sklon & imenovalnik \\
+živost & 0 \\
+vid & 0 \\
+oblika & 0 \\
+oseba & 0 \\
+nikalnost & 0 \\
+stopnja & nedoločeno \\
+določnost & ne \\
+število\_svojine & 0 \\
+spol\_svojine & 0 \\
+naslonskost & 0 \\
+zapis & 0 \\
+\end{tabular}
+\end{center}
+\caption{Whole MSD for \textit{Lep} in Slovene}
+\label{tab:MSDLep}
+\end{table}
+
 The first output was creted with function \textit{prettyPrintWithDescription}, full output can be created with \textit{prettyPrintWithFullDescription}. Both function are defined in \texttt{example.py}
@@
 Solution is described in the following section.
+
 \subsection{Encoding problems} %BANIČ
-
 Printing the result of our example is not optimal.
@@
 \par
-We also contacted the NLTK developers group and will arrange our taggers to become the part of NLTK.
+Our whole project is publicly available on github \cite{CODE}. We also contacted the NLTK developers group and will arrange our taggers to become the part of NLTK.
 
@@
 \bibitem{tokenizer} \url{http://docs.huihoo.com/nltk/0.9.5/guides/tokenize.html}
+\bibitem{CODE} \url{https://github.com/nikicc/slovene-nltk-tagger}
 \end{thebibliography}

### 2012-01-08 paper/paper.tex
Fixed some errors.

@@
 First, let us say a few words about the structure of the paper.
-In this chapter, we provide a breif theorethical description of imlemented taggers.
+In this chapter, we provide a brief theoretical description of implemented taggers.
 Next we will describe the related work in section \ref{relatedWork}, which was the base of our project.
-In section \ref{implementation} the proces of implementaion itself is explained.
+In section \ref{implementation} the process of implementation itself is explained.
 The example, how to use the tagger is provided, along with some discussion on encoding problems.
@@
 %NIKICC
-\subsection[Trigram tagger]{Trigram tagger\footnote{Definition and images summerized from \cite{NLTKBOOK}}}
-To understant how \textit{trigram} works, we should first understand \textit{unigram}.
+\subsection[Trigram tagger]{Trigram tagger\footnote{Definition and images summarized from \cite{NLTKBOOK}}}
+To understand how \textit{trigram} works, we should first understand \textit{unigram}.
 Unigram taggers are based on a simple statistical algorithm: \textit{for each token, assign the tag that is most likely for that particular token}.
-In the case of tagging with unigram tagger, we only consider the current token, in isolation from any larger context. The best we can do is tag each word with its \textit{a priori} most likely tag.
+In the case of tagging with Unigram tagger, we only consider the current token, in isolation from any larger context. The best we can do is tag each word with its \textit{a priori} most likely tag.
 \par
-An n-gram tagger is a generalization of a unigram\footnote{Unigram could also be reffered to as 1-gram} tagger whose context is the current word together with the part-of-speech tags of the $n-1$ preceding tokens, as shown in figure \ref{fig:trigram}.
+An n-gram tagger is a generalization of a unigram\footnote{Unigram could also be refereed to as 1-gram} tagger whose context is the current word together with the part-of-speech tags of the $n-1$ preceding tokens, as shown in figure \ref{fig:trigram}.
 
@@
 \par
-Altrough we use the name trigram all throughtout this paper, we acctually refer to sequential tagger, which consists of
+Although we use the name trigram all throughout this paper, we actually refer to sequential tagger, which consists of
 \textit{affix}\footnote{The Affix tagger learns prefix and suffix patterns to determine the part of speech tag for word.}, \textit{unigram}, \textit{bigram} and \textit{trigram} taggers in this order respectively.
@@
 Despite obvious differences between Brill tagger and n-gram tagger there is one similarity.
-Both of them use \textit{supervised learning} method, since we require annotated training data to make sure whether the tagger's guess is a mistake or not.
+Both of them use \textit{supervised learning} method, since we require annotated training data to make sure whether the taggers guess is a mistake or not.
 
@@
 
-\subsection[Naive Bayes classifier]{Naive Bayes classifier\footnote{Definition and images summerized from \cite{NLTKBOOK}}}
+\subsection[Naive Bayes classifier]{Naive Bayes classifier\footnote{Definition and images summarized from \cite{NLTKBOOK}}}
 Both Brill and Trigram are sequential backoff taggers. Meanwhile, Naive Bayes is \textit{classifier based tagger}.
@@
 \item whether an email is spam or not.
-\item diagnosis to a given patient as described by observed characteristics of the patient
+\item diagnosis to a given patient as described by observed characteristics of the patient.
 \end{itemize}
@@
 Features make their contribution to the decision by voting against those labels, that don't occur with this feature.
-Likelihood score is therefore reduced for each label by multiplying it with the probability that an input value with that label would have that feature.
-Suppose that the word occurs ran occurs in  12\% of the sports documents, 10\% of the murder mystery documents, and 2\% of the automotive documents.
+Likelihood score is therefore reduced for each label by multiplying it with the probability, that an input value with that label would have that feature.
+Suppose that the word \textit{ran} occurs in  12\% of the sports documents, 10\% of the murder mystery documents, and 2\% of the automotive documents.
 The likelihood score for the sports label will be multiplied by 0.12, the likelihood score for the murder mystery label will be multiplied by 0.1, and the likelihood score for the automotive label will be multiplied by 0.02.
-The overall effect will be to slightly reduce the score of the sports label, reduce the murder mystery label (more then sports label) and significantly reduce the automotive label.
+The overall effect will be, to slightly reduce the score of the sports label, reduce the murder mystery label (more then sports label) and significantly reduce the automotive label.
 Figure \ref{fig:naive-bayes-bargraph} illustrates this example.
@@
 Its an open-source project hosting on Github, with moto to \textit{Train NLTK objects with zero code}.
-We almost exclusively used the script \texttt{train\_tagger.py}, which has the ability to generate NLTK taggers. This project made it a lot easier for us to train the taggers on Slovenian corpuses.
+We almost exclusively used the script \texttt{train\_tagger.py}, which has the ability to generate NLTK taggers. This project made it a lot easier for us to train the taggers on Slovenian corpora.
 
@@
 It contains collections of various text in Slovenian language with annotation.
-For our project we used jos1M corpus that contain 1 million words with it's appropriate lemmas and morphosyntactic descriptions.
+For our project we used jos1M corpus, that contain 1 million words with it's appropriate lemmas and morphosyntactic descriptions.
 The JOS is compatible with Slovene\textit{ MULTEXT-East morphosyntactic specifications Version 4}\cite{MULTEXT-East}.
-It basic purpose is to define word classes.
+It basic purpose is to define word classes (i.e. part-of-speech).
 For each class there are various attributes and their appropriate values, which they can be mapped into morphosyntactic descriptions (i.e. MSD).
@@
 We can easily manipulate XML files in various programming languages (e.g. Python with xml.dom.minidom).
-This is mandatory because \textit{nltk-trainer}\cite{nltk-trainer} expects special form of input file (.pos file).
+This is mandatory, because \textit{nltk-trainer}\cite{nltk-trainer} expects special form of input file (.pos file).
 
@@
 
-The trainer will train the tagger properly, if in .pos file:
+The trainer will train the tagger properly, if .pos file consists off:
 \begin{itemize}
@@
 Beside this, there are many optional arguments.
-Here we will describe basics, for further description run script help with
+Here we will describe basics, for further description run script help with:
 \begin{lstlisting}
@@
 \par
-One is wheather to generate a Sequential Tagger, a Brill Tagger or a Classifier Based Tagger.
+One is whether to generate a Sequential Tagger, a Brill Tagger or a Classifier Based Tagger.
 For Sequential we can then select sequential backoff algorithm.
@@
 The default is \textit{aubt}, which was used for our Trigram tagger, but you can set this to the empty string to not train a sequential backoff tagger.
-The Brill Tagger is trained in front of the other tagger, which in our implementation was the Trigram tagger.
+The Brill tagger is trained in front of the other tagger, which in our implementation was the Trigram tagger.
 When building classifier based algorithm, we must select the classifier or the sequence of classifier.
@@
 \par
-Last argument, that should be mentioned, is wheather to evaluate the tagger or not and the procentage of sentaces to be used.
-The default value is 1, which means that the \textit{same} sentances are first used for training the tagger and then for evaluation.
+Last argument, that should be mentioned, is whether to evaluate the tagger or not and the percentage of sentences to be used.
+The default value is 1, which means that the \textit{same} sentences are first used for training the tagger and then for evaluation.
 We recommend this argument to be set below 1.
 If so, the right approach is used.
-Sentances are devided into two groups, one in used for training and second for evaluation. \textit{Fraction} is the procentage of sentances assigned to training group. 
+Sentences are divided into two groups, one in used for training and second for evaluation. \textit{Fraction} is the percentage of sentences assigned to training group. 
 This was used for evaluating constructed taggers. For results of this evaluation see section \ref{results}).
@@
 
-As you can see, there is the problem with the last dot. The solution we used is described below.
+As you can see, there is a problem with the last token. The solution we used is described below.
 
@@
 
-We also implemented a function that returns two description dictionaries for MSDs with attribute=value expansions in Slovene and English.
+We also implemented a function that returns two description dictionaries for MSD's with attribute=value expansions in Slovene and English.
 This way user can view which part-of-speech words belong to in Slovene or English.
@@
 
-The first output was creted with function \textit{prettyPrintWithDescription}, full output can be created with \textit{prettyPrintWithFullDescription}. Both function are defined in \texttt{example.py}
+The first output was created with function \textit{prettyPrintWithDescription}, full output can be created with \textit{prettyPrintWithFullDescription}.
+Both function are defined in \texttt{example.py}
 \par
@@
 For all constructed taggers, we here provide their accuracy.
-The results were obtained by repeating the proces of building the tagger with various evaluation arguments (see script \texttt{evaluateTaggers.sh}).
+The results were obtained by repeating the process of building the tagger with various evaluation arguments (see script \texttt{evaluateTaggers.sh}).
 See table \ref{tab:evaluation} and figure \ref{fig:evaluation}.
@@
 Under the assumption that most of the time was actually spent for tagging the sentences (besides just testing for equality and percentage calculation are needed), we noticed that the time for evaluations differ drastically.
-The results of resarching time consumption are listed in table \ref{tab:evaluation_speed} and graphically represented on figure \ref{fig:evaluation_speed} (see script \texttt{evaluateTaggersSpeed.py}).
+The results of researching time consumption are listed in table \ref{tab:evaluation_speed} and graphically represented on figure \ref{fig:evaluation_speed} (see script \texttt{evaluateTaggersSpeed.py}).
 
@@
 
-The difference here is enormous. Brill and Trigram run at almost the same speed, while Naive Bayes runs extremely slower. This is also one of the reason why we prefer usig the Brill tagger.
+The difference here is enormous. Brill and Trigram run at almost the same speed, while Naive Bayes runs extremely slower. This is also one of the reason why we prefer using the Brill tagger.
 
@@
 To conclude our paper, tagging in Slovene language is practical with proper use of Natural Language Toolkit and \textit{nltk-trainer}.
-Time spent in seconds for tagging different numbers of words, show that Brill and Trigram ran at the same speed, while Naive Bayes ran drasticly slower.
+Time spent in seconds for tagging different numbers of words, show that Brill and Trigram ran at the same speed, while Naive Bayes ran drastically slower.
 There are no major differences in tagging accuracy.
@@
 The main reason we built only these three is computing power.
-These three are the only one we could built and evaluate in some resonable time.
+These three are the only one we could built and evaluate in some reasonable time.
 \par

### 2012-01-12 example.py
Removed one print in example.py.

@@ -86,7 +86,6 @@ uText = 'Lep je dan, vse diši že po pomladi!'
 
 _t2 = time.time()
 result = sloTag2(uText)
-print result
 _t3 = time.time()
 
 prettyPrintWithDescription(result, dict_sl)

### 2012-01-15 paper/paper.tex
Fixing errors for final version

@@
 Tagger is an object, which processes a sequence of words and attaches a part of speech tag to each word.
-Our language of implemetation was Slovene.
+Our implementation language was Slovene.
 We constructed sequential \textit{Trigram tagger}, \textit{Brill tagger} and classifier based \textit{Naive Bayes tagger}, which differ in tagging accuracy and time consumed for tagging.
@@
 \section{Introduction} %NIKICC
-First, let us say a few words about the structure of the paper.
+\textit{Part-of-speech tagging} is harder than just having a list of words and their parts of speech, because some words can represent more than one part of speech at different times and because some parts of speech are complex or unspoken.
+This is not rare in natural languages (as opposed to many artificial languages), a large percentage of word-forms are ambiguous.
+Native speakers of a language perform grammatical and semantic analysis innately, and thus trained linguists can identify the grammatical parts of speech to various fine degrees depending on the tagging system \cite{wiki}.
+Although the problem is not trivial for a computer, the taggs are useful categories for many language processing tasks.
+\par
 In this chapter, we provide a brief theoretical description of implemented taggers.
-Next we will describe the related work in section \ref{relatedWork}, which was the base of our project.
+Next we will describe the tools used in section \ref{toolsUsed}, which were the base of our project.
 In section \ref{implementation} the process of implementation itself is explained.
@@
 %NIKICC
-\subsection[Trigram tagger]{Trigram tagger\footnote{Definition and images summarized from \cite{NLTKBOOK}}}
+\subsection{Trigram tagger}
 To understand how \textit{trigram} works, we should first understand \textit{unigram}.
@@
 \par
-An n-gram tagger is a generalization of a unigram\footnote{Unigram could also be refereed to as 1-gram} tagger whose context is the current word together with the part-of-speech tags of the $n-1$ preceding tokens, as shown in figure \ref{fig:trigram}.
+An n-gram tagger is a generalization of a unigram tagger whose context is the current word together with the part-of-speech tags of the $n-1$ preceding tokens, as shown in figure \ref{fig:trigram}.
+Unigram could also be refereed to as 1-gram.
 
@@
 In the example of an n-gram tagger shown in figure \ref{fig:trigram}, we have $n=3$; that is, we consider the tags of the two preceding words in addition to the current word.
-An n-gram tagger picks the tag that is most likely in the given context.
+An n-gram tagger picks a priori most likely tag in the given context.
 \par
-Although we use the name trigram all throughout this paper, we actually refer to sequential tagger, which consists of
-\textit{affix}\footnote{The Affix tagger learns prefix and suffix patterns to determine the part of speech tag for word.}, \textit{unigram}, \textit{bigram} and \textit{trigram} taggers in this order respectively.
-First we build affix, which is used as ad \textit{backoff-tagger} for unigram.
+Although we use the name trigram throughout this paper, we actually refer to sequential tagger, which consists of
+\textit{affix}, \textit{unigram}, \textit{bigram} and \textit{trigram} taggers in this order respectively.
+The Affix tagger learns prefix and suffix patterns to determine the part of speech tag for a word.
+First we build affix, which is used as a \textit{backoff-tagger} for unigram.
 Unigram is afterwards used as a backoff-tagger for bigram, which is finally used for trigram.
+\par
+Definitions and images in this section are summarized from \cite{NLTKBOOK}.
 
@@
 
-\section{Related work}
-\label{relatedWork}
+\section{Tools used}
+\label{toolsUsed}
 \subsection{Nltk-trainer} %NIKICC
@@
 \subsection{Training the tagger} %NIKICC
-As mentioned above, the script \texttt{train\_tagger.py} was used for this purpose.
-Its mandatory argument is a corpus in .pos format, which generation is described in \ref{Corpus transformation}.
-Beside this, there are many optional arguments.
-Here we will describe basics, for further description run script help with:
+The script \texttt{train\_tagger.py} was used for this purpose.
+Its mandatory argument is a corpus in .pos format.
+There are many optional arguments.
+We will describe only the basic ones, for further description run script help with:
 \begin{lstlisting}
@@
 \par
-One is whether to generate a Sequential Tagger, a Brill Tagger or a Classifier Based Tagger.
-For Sequential we can then select sequential backoff algorithm.
-This can be any combination of the following letters:
+To generate a Sequential Tagger we use:
 \begin{itemize}
@@
 \end{itemize}
-The default is \textit{aubt}, which was used for our Trigram tagger, but you can set this to the empty string to not train a sequential backoff tagger.
-The Brill tagger is trained in front of the other tagger, which in our implementation was the Trigram tagger.
-When building classifier based algorithm, we must select the classifier or the sequence of classifier.
+We can use any combination of the these letters, to select a sequential backoff algorithm.
+The default is \textit{aubt}, which was used for our Trigram tagger, but you can set this to the empty string not to train a sequential backoff tagger.
+The Brill tagger is trained before the other taggers, which in our implementation was the Trigram tagger.
+When building a classifier based algorithm, we must select the classifier or the sequence of classifier.
 Following are supported: NaiveBayes, DecisionTree, Maxent, GIS, IIS, CG, BFGS, Powell, LBFGSB, Nelder-Mead, MEGAM and TADM.
-We should mention, that the time spent for training classifier based tagger, is much larger than, the time for a sequential or brill.
+The time to train classifier based tagger, is much larger than, the time for a sequential or Brill.
 \par
-Script also has the possibility to set default tag. We used \textit{-Neznan-}, which is Slovene word for \textit{-unknown-}.
+Script also has the possibility to set a default tag. We used \textit{-Neznan-}, which is Slovene word for \textit{-unknown-}.
 \par
-Last argument, that should be mentioned, is whether to evaluate the tagger or not and the percentage of sentences to be used.
+We also choose whether to evaluate the tagger or not and the percentage of sentences to be used.
 The default value is 1, which means that the \textit{same} sentences are first used for training the tagger and then for evaluation.
-We recommend this argument to be set below 1.
-If so, the right approach is used.
-Sentences are divided into two groups, one in used for training and second for evaluation. \textit{Fraction} is the percentage of sentences assigned to training group. 
-This was used for evaluating constructed taggers. For results of this evaluation see section \ref{results}).
+We recommend that this argument is <1.
+In this case, sentences are divided into two groups.
+The first in used for training and the second for evaluation.
+\textit{Fraction} is the percentage of sentences assigned to training group.
+This was used for evaluating constructed taggers. For results of this evaluation see section \ref{results}.
 
@@
 \label{results}
-For all constructed taggers, we here provide their accuracy.
+For all constructed taggers, we provide their accuracy.
 The results were obtained by repeating the process of building the tagger with various evaluation arguments (see script \texttt{evaluateTaggers.sh}).
-See table \ref{tab:evaluation} and figure \ref{fig:evaluation}.
+See Table \ref{tab:evaluation} and Figure \ref{fig:evaluation}.
 
@@
 \end{center}
-\caption{Taggers accuracy for various fractions.}
+\caption{Taggers accuracy for various values of fraction parameter.}
 \label{tab:evaluation}
@@
 \par
-To conclude, there are no major differences in tagging accuracy. The Brill and the Naive Bayes are about 2\% better than Trigram.
+As seen, there are no major differences in tagging accuracy. The Brill and the Naive Bayes are about 2\% better than Trigram.
 \par
-However, during this evaluation we discover an other parameter, which should be taken into account.
-Under the assumption that most of the time was actually spent for tagging the sentences (besides just testing for equality and percentage calculation are needed), we noticed that the time for evaluations differ drastically.
-The results of researching time consumption are listed in table \ref{tab:evaluation_speed} and graphically represented on figure \ref{fig:evaluation_speed} (see script \texttt{evaluateTaggersSpeed.py}).
+We noticed that the time for evaluations differ drastically.
+The results are listed in Table \ref{tab:evaluation_speed} and graphically represented on Figure \ref{fig:evaluation_speed} (see script \texttt{evaluateTaggersSpeed.py}).
 
@@
 
-The difference here is enormous. Brill and Trigram run at almost the same speed, while Naive Bayes runs extremely slower. This is also one of the reason why we prefer using the Brill tagger.
+The difference is enormous. Brill and Trigram run at almost the same speed, while Naive Bayes runs much slower. This is also one of the reason why we prefer using the Brill tagger.
 
@@
 
-\section*{Acknowledgements}
-We would like to express our gratitude to all the contributors of \textit{JOS} \cite{JOS} project, \textit{nltk-trainer} \cite{nltk-trainer} project and to Marko Robnik-Šikonja, PhD, Associate Professor, who was our mentor at Faculty of Computer and Information Science.
-Also this project would not be possible without \textit{Natural Language Toolkit} \cite{nltk} and their admirable documentation.
+\section*{Acknowledgements} %NIKICC
+We would like to express our gratitude to all the contributors of \textit{JOS} \cite{JOS} project, \textit{nltk-trainer} \cite{nltk-trainer} project and to Prof.
+Marko Robnik-Šikonja, who was our supervisor at Faculty of Computer and Information Science.
+Also this project would not be possible without \textit{Natural Language Toolkit} \cite{nltk} and its admirable documentation.
 
 \begin{thebibliography}{99}
-\bibitem{nltk} \url{http://www.nltk.org/}
-\bibitem{JOS} \url{http://nl.ijs.si/jos/}
-\bibitem{nltk-trainer} \url{https://github.com/japerk/nltk-trainer}
-\bibitem{MULTEXT-East} \url{http://nl.ijs.si/ME/V4/msd/html/msd-sl.html}
-\bibitem{NLTKBOOK} \url{http://www.nltk.org/book}
-\bibitem{xml_tags} \url{http://nl.ijs.si/ssj/schema/tei_ssj_doc.pdf}
-\bibitem{tokenizer} \url{http://docs.huihoo.com/nltk/0.9.5/guides/tokenize.html}
-\bibitem{CODE} \url{https://github.com/nikicc/slovene-nltk-tagger}
+\bibitem{nltk} (2011) NLTK web page. [Online]. Available at: \url{http://www.nltk.org/}
+\bibitem{JOS} (2011) JOS project web page. [Online]. Available at: \url{http://nl.ijs.si/jos/}
+\bibitem{nltk-trainer} (2011) NLTK-trainer project source code. [Online]. Available at: \url{https://github.com/japerk/nltk-trainer}
+\bibitem{MULTEXT-East} (2011) Slovene MULTEXT-East Morphosyntactic Specifications. [Online]. Available at: \url{http://nl.ijs.si/ME/V4/msd/html/msd-sl.html}
+\bibitem{NLTKBOOK} S.
+Bird, E.
+Klein, and E.
+Loper.
+(2009, June).
+Natural Language Processing with Python \textit{--- Analyzing Text with the Natural Language Toolkit}.
+(1st ed.) [Online].
+Available at: \url{http://www.nltk.org/book}
+\bibitem{xml_tags} T. Erjavec. (2011, January 29). TEI Schema for SSJ and JOS corpora of Slovene. [Online].
+Available at: \url{http://nl.ijs.si/ssj/schema/tei_ssj_doc.pdf}
+\bibitem{tokenizer} (2011) Tokenizers. [Online]. Available at: \url{http://docs.huihoo.com/nltk/0.9.5/guides/tokenize.html}
+\bibitem{CODE} (2011) Slovene NLTK Tagger (the source code from our project). [Online]. Available at: \url{https://github.com/nikicc/slovene-nltk-tagger}
+\bibitem{wiki} (2011) Wikipedia. [Online]. Available at: \url{http://en.wikipedia.org/wiki/Part-of-speech_tagging}
 \end{thebibliography}

### 2012-01-15 paper/paper.tex
Corrected some mistakes.

@@
 
-The Brill tagger is another part-of-speech tagger and was first described by Eric Brill in 1993. Before trying to explain it, we must address issues with n-gram taggers (e.g. \textit{Trigram tagger}).
+First problem with n-gram taggers is the size of their n-gram table. If we want to use tagging process on mobile computing device, compromise must be taken between performance and model size.
 
-First issue with n-gram taggers is size of their n-gram table. If we want to use tagging process on mobile computing device, compromise must be taken between performance and model size.
+Second problem with n-gram tagger is, that the only information it considers from the prior context are tags.
+This is not practical because words themselves are useful information.
+Therefore n-gram taggers are impractical models for identification of words in context.
 
-A second issue with n-gram tagger is, that the only information it considers from prior context is tags.
-This is not practical because words themselves are useful information also.
-Therefore n-gram taggers are impractical models to be used to identify words in context.
-
-To bypass those issues we can use \textit{Brill tagger}.
+To bypass those problems we can use \textit{Brill tagger}.
+The Brill tagger was first described by Eric Brill in 1993.
 It can be summarised as an transformation-based learning and performs very well using models that are just a fraction of the size of n-gram taggers.
@@
 Despite obvious differences between Brill tagger and n-gram tagger there is one similarity.
-Both of them use \textit{supervised learning} method, since we require annotated training data to make sure whether the taggers guess is a mistake or not.
+Both of them use \textit{supervised learning} method, since we require annotated training data to check whether the taggers guess is a mistake or not.
 
@@
 
-\subsection[Naive Bayes classifier]{Naive Bayes classifier\footnote{Definition and images summarized from \cite{NLTKBOOK}}}
-Both Brill and Trigram are sequential backoff taggers. Meanwhile, Naive Bayes is \textit{classifier based tagger}.
-Classification is a process where, for a given output, we have to choose the correct class. Basically, it is an  procedure for assigning input data into one of a given number of categories.
+\subsection{Naive Bayes classifier}
+Both Brill and Trigram are sequential backoff taggers. Meanwhile, Naive Bayes is a \textit{classifier based tagger}.
+Classification is a process where, for a given output, we have to choose the correct class. Basically, it is  procedure for assigning input data into one of a given categories.
 
@@
 \begin{itemize}
-\item whether an email is spam or not.
-\item diagnosis to a given patient as described by observed characteristics of the patient.
+\item is an email a spam or not.
+\item diagnosis of a given disease as described by observed characteristics of the patient.
 \end{itemize}
@@
 Later it must take into consideration the effect of each feature.
-Example considers, that input document contains word "dark", which indicates for "murder mystery", but also contains the word "football", which (strongly) indicate for "sports".
+Example considers, that input document contains word "dark", which indicates to "murder mystery", but also contains the word "football", which (strongly) indicates to "sports".
 After every feature has made its contribution, the classifier checks which label it is closest to, and assigns that label to the input.
@@
 
+Definitions and images in this section are summarized from \cite{NLTKBOOK}.
 \section{Tools used}
@@
 It stands for: \textit{Jezikoslovno Označevanje Slovenskega jezika} (Linguistic annotation of Slovenian language).
-It contains collections of various text in Slovenian language with annotation.
-For our project we used jos1M corpus, that contain 1 million words with it's appropriate lemmas and morphosyntactic descriptions.
+It contains collections of various text in Slovenian language with annotations.
+For our project we used jos1M corpus,  containing 1 million words with appropriate lemmas and morphosyntactic descriptions.
 The JOS is compatible with Slovene\textit{ MULTEXT-East morphosyntactic specifications Version 4}\cite{MULTEXT-East}.
-It basic purpose is to define word classes (i.e. part-of-speech).
-For each class there are various attributes and their appropriate values, which they can be mapped into morphosyntactic descriptions (i.e. MSD).
-The structure of JOS corpora is in  \textit{Extensible Markup Language (XML)}.
+Its basic purpose is to define word classes (i.e. part-of-speech).
+For each class there are various attributes and their appropriate values, which can be mapped into morphosyntactic descriptions (i.e. MSD).
+The structure of JOS corpora is described in a form of \textit{Extensible Markup Language (XML)}.
 
@@
 \item Every punctuation mark is separated with slash and then follows the same punctuation mark.
-\item Words and punctuation marks are between space.
+\item Words and punctuation marks are separated with spaces.
 \end{itemize}
@@
 If we follow these "rules", implementation is simple.
-We just need to go through all child tags in sentence tag and properly write them in a file.
+We go through all child tags in a sentence tag and write them to a file.
 Because Slovenian language contain letters with caron, proper encoding of tags must be provided.
-This is simply done with function \textit{encode()} (e.g.
-encode("utf-8") - this is proper encoding for Slovenian language).
-\textit{Jos1M} corpus is build up with 10 XML files (e.g. \texttt{jos1M-01.xml},  \texttt{jos1M-02.xml}, ..., \texttt{jos1M-10.xml}), so we must parse all of them. 
+This is done with function \textit{encode()} (e.g.
+encode("utf-8") - for Slovenian language).
+\textit{Jos1M} corpus is build from 10 XML files (e.g. \texttt{jos1M-01.xml},  \texttt{jos1M-02.xml}, ..., \texttt{jos1M-10.xml}), so we parse all of them. 
 File \texttt{transformJOS.py} contains the corpus transformation.
@@
 
-\subsection{Usage with NLTK} %BANIČ
+\subsection{Use of taggers with NLTK} %BANIČ
 
-We can use the trained tagger and NLTK to see how they can be used in Slovenian language. The requirements are installed NLTK and Python 2.7. We set an example in file \texttt{example.py}.
-First we need to import following functions:
+We can apply the trained tagger and NLTK to see how they can be used in Slovenian language. The pre-requirements are NLTK and Python 2.7. We present an example in file \texttt{example.py}.
+First we need to import the functions:
 \begin{itemize}
@@
 These tokens are then used by the tagger.
-Following example will demonstrate, how does PunktWordTokenizer tokenize a sentence.
 If we have the following sentence:
@@
 [\textit{'Potem', 'se', 'je', 'obrnil', 'proti', 'nam', 'in', 'skozi', 'solze', 'izoblikoval', 'najlepši', 'nasmeh', ',', 'kar', 'sem', 'jih', 'videl', 'v', 'življenju.'}]
+\par
 
-As you can see, there is a problem with the last token. The solution we used is described below.
+As you can see, there is a problem with the last token, because the tokenizer did not separate \textit{življenju} and period.
+The tagger will not recognize [\textit{življenju.}], because it is not in a valid form (i.e.
+words and punctuation marks must be separated tokens).
+Before tagging we must separate this type of tokens into two new tokens.
 
@@
 We use it for \textit{sentence segmentation}.
-In some cases user can input not a sentence, but a whole paragraph.
+In some cases the user can input a whole paragraph.
 Before using PunktWordTokenizer to tokenize sentence into tokens, we need to segment it into sentences.
-This can be easily used by including Slovene Punkt sentence segmenter.
-Following example shows, how does sentence segmenter work.
-If we have the following paragraph:
+This can be easily done by including Slovene Punkt sentence segmenter.
+The example shows, how sentence segmenter works:
 
@@
 
-the sentence segmenter will output the following list of sentences:
+The sentence segmenter will output the following list of sentences:
 
@@
 The first step is to use sentence segmenter, to form a list of sentences.
-This list is then being tokenized with PunktWordTokenizer.
-One drawback with this tokenizer is that it doesn't tokenize the last word of a sentence and the following period (as shown in the example).
+This list is then tokenized with PunktWordTokenizer.
+A drawback is that the last word of a sentence and the following period (as shown in the example) are not tokenized.
 Before using the PunktWordTokenizer, we must divide them with space.
 After we tokenize the sentences, tagging can be used.
-The result is a list of tuples with words or characters and their appropriate morphosyntactic description.
+The result is a tuple list with words or characters and their morphosyntactic description.
 
 We also implemented a function that returns two description dictionaries for MSD's with attribute=value expansions in Slovene and English.
-This way user can view which part-of-speech words belong to in Slovene or English.
-Following example demonstrates the expected output:
+Thereby a user can check which part-of-speech words belong to Slovene or English.
+The expected output is:
 \\\\
@@
 
-We must exmplain, that here only the part of whole MSD is printed. For example the whole MSD for \textit{Lep} shown in table \ref{tab:MSDLep}.
+Here only a part of the whole MSD is presented. The whole MSD for \textit{Lep} is shown in table \ref{tab:MSDLep}.
 
@@
 The first output was created with function \textit{prettyPrintWithDescription}, full output can be created with \textit{prettyPrintWithFullDescription}.
-Both function are defined in \texttt{example.py}
+Both function are defined in \texttt{example.py}.
 \par
-But there is an issue.
-Because Slovenian language contain letters with caron, the output will have unicode symbols for them.
-In the previous examples, the output is already corrected.
-Solution is described in the following section.
+Because Slovenian language contain letters with caron, the output will have unicode symbols for them. In the previous examples, the output is already corrected.
 
-\subsection{Encoding problems} %BANIČ
 Printing the result of our example is not optimal.
-Because we want users to know what these characters are, we implemented a solution.
-To resolve this, we must decode every word to \textit{UTF - 8} (because words we tag are in UTF - 8).
+We decode every word to \textit{UTF - 8} (because words we tag are in UTF - 8).
 If a user wants to print the result in "pretty" form, he should call function \textit{prettyPrint()} or \textit{prettyPrintByLine()} (both are implemented in \texttt{example.py}).
@@
 To conclude our paper, tagging in Slovene language is practical with proper use of Natural Language Toolkit and \textit{nltk-trainer}.
-Time spent in seconds for tagging different numbers of words, show that Brill and Trigram ran at the same speed, while Naive Bayes ran drastically slower.
-There are no major differences in tagging accuracy.
-However Brill and the Naive Bayes are about 2\% better than Trigram.
+Time for tagging show that Brill and Trigram ran at the same speed, while Naive Bayes is much slower.
+There are no major differences in tagging accuracy, but Brill and the Naive Bayes are about 2\% better than Trigram.
 \par
-Besides these three taggers we built, others can be built without any bigger modification.
-All that is needed is to set some arguments when calling the script \texttt{train\_tagger.py}.
-The main reason we built only these three is computing power.
-These three are the only one we could built and evaluate in some reasonable time.
+Besides the taggers we built, others can be built without any large modifications by calling the script \texttt{train\_tagger.py}.
 \par
-Our whole project is publicly available on github \cite{CODE}. We also contacted the NLTK developers group and will arrange our taggers to become the part of NLTK.
+Our whole project is publicly available on github \cite{CODE}. We contacted the NLTK developers group and will arrange our taggers to become a part of NLTK.
 

### 2012-01-15 paper/paper.tex
Improved the conclustion.

@@
 Besides the taggers we built, others can be built without any large modifications by calling the script \texttt{train\_tagger.py}.
+We have no proof that the ones we build are better than the ones we did not.
+It is possible, that some other tagger would be more accurate or faster or even both.
+So here is definitely some room for improvements.
+\par
+If we would like to make the taggers more accurate, including other corpuses into the proces of learning would probably help.
+All we need to do is transform new corpuses into .pos format.
+The Slovene language has more complicated structure than English, so the corpus needed should be larger.
 \par

### 2012-01-15 paper/paper.tex
Added Brill algorithm.

@@
 
+The algorithm starts with the assignment of tags in a maximum likelihood manner; then improving "patches" are determined via the generation of rules that correct tagging errors made in the initialization phase \cite{brill}:
+\begin{itemize}
+\item[•]Initialization:
+\begin{itemize}
+\item[•]Known words (in vocabulary): assigning the most frequent tag associated to a form of the word
+\item[•]Unknown words (out of vocabulary) :
+\begin{itemize}
+\item[•] Proper noun if capitalized and simple noun else
+\item[•] Learning or guessing rules with lexical rules on the same basis as contextual rules
+\end{itemize}
+\end{itemize}
+
+\item[•] Learning Phase:
+\begin{itemize}
+\item[•] Iteratively compute the error score of each candidate rule (difference between the number of errors before and after applying the rule)
+\item[•] Select the best (higher score) rule.
+\item[•] Add it to the rule set and apply it to the text.
+\item[•] Repeat until no rule has a score above a given threshold (that is, if the chosen threshold is zero (which can lead to over-fitting), until applying new rules leaves the text in the same state, which is then supposed to be the final state of the tagging).
+\end{itemize}
+\end{itemize}
+
+Lexical rules are used for the initialization, and contextual rules are used to correct the tags.
+\begin{itemize}
+\item[•] Lexical rules: word to tag IF Condition (example: identification of suffixes like "-tion")
+\item[•] Contextual rules: tag1 to tag2 IF Condition (example: "preceding/following tag is X", "preceding/following word is w")       
+\end{itemize}
 \subsection{Naive Bayes classifier}
@@
 \bibitem{wiki} (2011) Wikipedia. [Online]. Available at: \url{http://en.wikipedia.org/wiki/Part-of-speech_tagging}
+\bibitem{brill} (2011) Brill tagger algorithm. [Online]. Available at: \url{http://en.wikipedia.org/wiki/Brill_tagger}
 \end{thebibliography}

### 2012-01-15 paper/paper.tex
Fixed some final errors.

@@
 Native speakers of a language perform grammatical and semantic analysis innately, and thus trained linguists can identify the grammatical parts of speech to various fine degrees depending on the tagging system \cite{wiki}.
-Although the problem is not trivial for a computer, the taggs are useful categories for many language processing tasks.
+Although the problem is not trivial for a computer, the tags are useful categories for many language processing tasks.
 \par
@@
 \par
-An n-gram tagger is a generalization of a unigram tagger whose context is the current word together with the part-of-speech tags of the $n-1$ preceding tokens, as shown in figure \ref{fig:trigram}.
+An n-gram tagger is a generalization of a unigram tagger whose context is the current word together with the part-of-speech tags of the $n-1$ preceding tokens, as shown in Figure \ref{fig:trigram}.
 Unigram could also be refereed to as 1-gram.
@@
 The tag to be chosen, $t_{n}$, is circled, and the context is shaded in grey.
-In the example of an n-gram tagger shown in figure \ref{fig:trigram}, we have $n=3$; that is, we consider the tags of the two preceding words in addition to the current word.
+In the example of an n-gram tagger shown in Figure \ref{fig:trigram}, we have $n=3$; that is, we consider the tags of the two preceding words in addition to the current word.
 An n-gram tagger picks a priori most likely tag in the given context.
@@
 The default is \textit{aubt}, which was used for our Trigram tagger, but you can set this to the empty string not to train a sequential backoff tagger.
+\par
 The Brill tagger is trained before the other taggers, which in our implementation was the Trigram tagger.
-When building a classifier based algorithm, we must select the classifier or the sequence of classifier.
+\par
+When building a classifier based algorithm, we must select the classifier or the sequence of classifiers.
 Following are supported: NaiveBayes, DecisionTree, Maxent, GIS, IIS, CG, BFGS, Powell, LBFGSB, Nelder-Mead, MEGAM and TADM.
@@
 It is possible, that some other tagger would be more accurate or faster or even both.
-So here is definitely some room for improvements.
+Here is definitely some room for improvements.
 \par
@@
 All we need to do is transform new corpuses into .pos format.
-The Slovene language has more complicated structure than English, so the corpus needed should be larger.
+The Slovene language has more complicated structure than English, so the learning corpus should be larger.
 \par

### 2012-01-16 slides/slides.tex
Strat making slides 2

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usetheme{Warsaw}
\title[Slovenski NLTK označevalnik]{Slovenski NLTK označevalnik}
\author{
	Niko Colnerič
	\and
	Nejc Banič}
\institute{ Fakulteta za Računalništvo in Informtiko\\
			Univerza v Ljubljani}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Motivacija}
This is a short introduction to Beamer class.
\end{frame}

\begin{frame}{Implementirani označevalniki}
\begin{itemize}
\item \textit{Trigram tagger}
\item \textit{Brill tagger}
\item classifier based \textit{Naive Bayes tagger}
\end{itemize}
\end{frame}

\begin{frame}{JOS korpus}
\end{frame}

\begin{frame}{NLTK trainer}
\end{frame}

\begin{frame}{Postopek treniranja označevalnikov}
\begin{enumerate}
\item Transformacija \textit{.XML} korpusa v \textit{.pos}
\item Treniranje z skripto iz NLTK-trainer
\item Ocenjevanje točnosti
\item Ocenjevanje hitrosti
\end{enumerate}
\end{frame}

\begin{frame}{Primer}
\begin{center}
\textit{\textbf{Lep je dan, vse diši že po pomladi}}
\end{center}
\textit{( Lep  |  PPNMEIN ) - pridevnik\\
	( je  |  GP-STE-N ) - glagol\\
	( dan  |  SOMEI ) - samostalnik\\
	( ,  |  , ) - ni razlage\\
	( vse  |  ZC-SEI ) - zaimek\\
	( diši  |  GGNSTE ) - glagol\\
	( že  |  L ) - členek\\
	( po  |  DM ) - predlog\\
	( pomladi  |  SOZEM ) - samostalnik\\
	( !  |  ! ) - ni razlage}\\
\end{frame}

\begin{frame}{Celoten MSD za \textit{Lep} }
\begin{table}[h]
\begin{center}
\begin{tabular}{c|c}
\multicolumn{2}{c}{pridevnik}\\\hline\hline
vrsta & splošni \\
spol & moški \\
število & ednina \\
sklon & imenovalnik \\
živost & 0 \\
vid & 0 \\
oblika & 0 \\
oseba & 0 \\
nikalnost & 0 \\
stopnja & nedoločeno \\
določnost & ne \\
število\_svojine & 0 \\
spol\_svojine & 0 \\
naslonskost & 0 \\
zapis & 0 \\
\end{tabular}
\end{center}
\label{tab:MSDLep}
\end{table}
\end{frame}

\begin{frame}{Rezultati natančnost}
\begin{figure}[h]
\begin{center}
\includegraphics[height=0.85\textheight]{../evaluation/graph.pdf} 
\end{center}
\end{figure}
\end{frame}

\begin{frame}{Rezultati hitrost}
\begin{figure}[h]
\begin{center}
\includegraphics[height=0.85\textheight]{../evaluation/graph_speed.pdf} 
\end{center}
\end{figure}
\end{frame}

\begin{frame}{Zaključek in izboljšave}
\end{frame}

\begin{frame}{}
\end{frame}

\end{document}

### 2012-01-16 slides/slides.tex
Slides...
New slides.

@@
 \begin{itemize}
-\item \textit{Trigram tagger}
-\item \textit{Brill tagger}
-\item classifier based \textit{Naive Bayes tagger}
+\item \textbf{Trigram označevalnik}\\
+\textit{a priori najbolj verjetno oznako v kontekstu dolžine 3}
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../paper/tag-context.png}
+\end{center}
+
+\item \textbf{Brillov označevalnik}\\
+\textit{ugane oznako vsako besede, nato 
+popravi svoje napake tako, da uporabi seznam transformacijskih pravil}
+\item označevalnik na podlagi \textbf{naivnega Bayes-ovega klasifikatorja}\\
+\textit{predpostavi pogojno neodvisnost atributov pri danem razredu}
 \end{itemize}
@@
 \begin{frame}{JOS korpus}
+
+Osnovna komponenta za izgradnjo slovenskega NLTK označevalnik je JOS korpus.
+Vsebuje zbirko različnih besedil s podrobno ročno preverjenimi jezikoslovnimi oznakami (npr.
+leme, oblikoskladenjske oznake itd.).
+Korpus je kompatibilen z MULTEXT-East V4 oblikoskladenjskimi specifikacijami (ang.
+\textit{morphosyntactic descriptions (MSDs)}), ki definirajo kategorije (besedne vrste) za slovenski jezik, za vsako kategorijo pa tudi njene atribute in njihove vrednosti.
+Uporabila sva korpus \textit{jos1M}, ki vsebuje 1 milijon besed z delno ročno preverjenimi lemami in oblikoskladenjskimi oznakami.
+Struktura JOS korpusa je zapisana v XML.
 \end{frame}
@@
 \begin{frame}{Celoten MSD za \textit{Lep} }
-\begin{table}[h]
 \begin{center}
@@
 \end{center}
-\label{tab:MSDLep}
-\end{table}
 \end{frame}
@@
 
-\begin{frame}{Zaključek in izboljšave}
+\begin{frame}{Zaključek, izboljšave}
+
+Ne moreva trditi, da so uporabljeni označevalniki (Trigram, Brill in naivni Bayes) najhitrejši in najbolj natančni. Zato je ena možnost za izboljšavo uporaba drugih označevalnikov. 
+\par
+Če bi hotela izboljšati natančnost naših označevalnik, bi lahko v proces učenja uporabila tudi drugih korpuse.
+
+\par
+Označevanje v slovenskem jeziku je praktično z uporabo knjižnice \textit{Natural Language Toolkit} in projekta nltk-trainer.
+Evalvacija je pokazala, da sta tako Brill kot Trigram označila besede hitro, medtem ko je bil naivni Bayes-ov klasifikator veliko počasnejši.
+Razlike v natančnosti vseh treh označevalnikov praktično ni.
 \end{frame}
-
-\begin{frame}{}
-\end{frame}
-
 \end{document}

### 2012-01-16 slides/slides.tex
Working on slides

--- 2012-01-16 slides/slides.tex
@@
 \usetheme{Warsaw}
+
 \title[Slovenski NLTK označevalnik]{Slovenski NLTK označevalnik}
@@
 \begin{frame}{Motivacija}
-This is a short introduction to Beamer class.
+\begin{itemize}
+\item NLTK - Natural Language Toolkit (Python 2.7)
+\item Označevalnik (tagger) besedam pripiše oblikoslovne oznake
+\item Težave z dvoumnostjo
+\item Uporabnost za nadaljno procesiranje teksta
+\item Tega so nas učili v šoli $\ddot\smile$
+\end{itemize}
 \end{frame}
@@
 \begin{frame}{JOS korpus}
+\begin{itemize}
+\item korpus = velika in strukturirana zbirka besedil
+\item Jezikoslovno Označevanje Slovenskega jezika (IJS)
+\item Zbirko besedil s ročno preverjenimi jezikoslovnimi oznakami
+\item \textit{.XML} datoteka
+\item Uporabila sva \textit{jos1M} - 1 milijon označenih besed
+\item MULTEXT-East V4 specifikacija
+\item MSD (morphosyntactic descriptions) - definirajo kategorije oz. besedne vrste
+\item Namenjena spodbujanju razvoja jezikovnih tehnologij za slovenski jezik
+\end{itemize}
 \end{frame}
@@
 \begin{frame}{NLTK trainer}
+\begin{itemize}
+\item Uporabljen za treniranje NLTK objektov - označevanlikov
+\item \textit{Train NLTK objects with zero code}
+\item Dobro dokumentiran
+\item Prosto dostopen na Github-u
+\end{itemize}
 \end{frame}
@@
 \begin{center}
-\textit{\textbf{Lep je dan, vse diši že po pomladi}}
+\textit{\textbf{Lep je dan, vse diši že po pomladi!}}
 \end{center}
@@
 
-\begin{frame}{Zaključek in izboljšave}
-\end{frame}
+\begin{frame}{Zaključek, izboljšave}
+\begin{itemize}
+\item Brill in Trigram sta hitra, NaiveBayes veliko počasnejši
+\item Najbolj natančen Brill, najmanj Trigram
+\item $\Rightarrow$ Priporočama uporabo Brill označevalnika
+\item Vključitev v NLTK
+\item Dostopno na: \url{https://github.com/nikicc/slovene-nltk-tagger}
+\end{itemize}
 
-\begin{frame}{}
+\begin{itemize}
+\item Z lahkoto bi zgradili tudi druge označevalnike, ki bi lahko bili boljši ali hitrejši
+\item Vključitev drugih korpusov, bi bila koristna
+\end{itemize}
 \end{frame}
